### 2018-05-09 report/MDP.tex
Add current report

\documentclass[a4paper,11pt]{article}
\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[]{geometry}
\usepackage{inconsolata}
\usepackage{courier}
\usepackage{listings}
\usepackage{enumerate}
\usepackage{float}
\usepackage{answers}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{commath}
\usepackage[labelfont=bf]{caption}
\usepackage{mwe}
\usepackage[longnamesfirst]{natbib}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{eurosym}
\usepackage{siunitx}
\usepackage{ctable}
\usepackage{subfig}
\usepackage{dsfont}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}

\title{Maintenance Policies for a Multi-Unit System with Adjustable Production Rates}
\author{Mark van der Broek}
\date{\today}

\begin{document}

	\maketitle

\section{Introduction}	
We consider a multi-unit system that consists of $n$ identical units. Each unit deteriorates according to a discrete-time Markov chain with deterioration states $X = \{0, 1, \dots, L\}$. State $0$ represents the as-good-as-new state and state $L$ the failed state. The system also allows to adjust the production rate of each unit, which has an impact on the deterioration process of the unit. The units deteriorate independently of each other, and we can perfectly observe this deterioration level for each unit at any moment in time. An graphical overview of the order of events within a time epoch is presented in Figure 1.

\begin{figure}[H]
	\centering
	\begin{tikzpicture}
	\coordinate (t) at (0,0);
	\coordinate (X) at (1,0);
	\coordinate (R) at (2,0);
	\coordinate (M) at (3.5,0);
	\coordinate (U) at (5,0);
	\coordinate (UE) at (9,0);
	\coordinate (D) at (10,0);
	\coordinate (t+1) at (11,0);

	% Draw the time horizon
	\draw[thick, |-|] (t) -- (t+1);
	\draw (t) node[below=3pt] {$t$};
	\draw[thick, ->] (1,1) node[above, align=center] {Observe $X_t$} -- (X);
	\draw[thick, ->] (2,-1) node[below, align=center] {Maintenance \\ decision $r$} -- (R);
	\draw[thick, ->] (3.5, 1) node[above, align=center] {Perform \\ maintenance} -- (M);
	\draw[thick, ->] (5, -1) node[below, align=center] {Production \\ decision $u$} -- (U);
	\draw[thick, decorate, decoration={brace,raise=2pt, amplitude=3pt}] (6,0) -- (9, 0) node[sloped,midway, above=6pt] {Produce};
	\draw[thick, ->] (10, 1) node[above, align=center] {System deteriorates} -- (D);
	\draw (t+1) node[below=3pt] {$t+1$};
	\end{tikzpicture}
	\caption{Order of events for time epoch $t$.}
\end{figure}

For the analysis I will consider a system of two units, and if time permits we will relax this assumption to consider a system with more units. For the maintenance policy can perform maintenance or not with cost $c_\text{pm}$ if the unit has not failed yet and cost $c_\text{cm}$ if it has failed. The cost of preventive maintenance is assumed to be less than corrective maintenance cost ($c_\text{pm} < c_\text{cm}$). Furthermore, for any time period we do perform maintenance a fixed cost $C$ has to be incurred. Therefore, depending on the size of these three costs it could be beneficial to cluster the maintenance actions.

\section{Markov Decision Process}

The system is modelled using a Markov Decision Process (MDP) with an infinite horizon. The system consists of $n$ identical units, each with $L+1$ deterioration states. We denote the set of units as $\mathcal{I} = \{1, \dots, n\}$. For unit $i \in \mathcal{I}$ we define the state as $x_i \in X$. The whole state of all units is denoted by
$$
x = (x_1, \dots, x_n).
$$

For each unit the set of actions at each decision epoch is to decide on the production rate, and to do no maintenance (DN) or to perform maintenance (M). The action space for the maintenance decision can be defined as
$$
R = \left \{r = (r_1, \dots, r_n): r_i \in \{\text{DN}, \text{M}\}, \enskip \forall i \in \mathcal{I} \right \}.
$$
Each unit can produce at most $m+1$ different production rates, ranging from $0$ to $1$ (maximum production). The possible production rates are uniformly distributed over the interval, i.e. $u_i \in \{j / m \mid j = 0, 1, \dots, m\}$. We let $u = (u_1, \dots, u_n)$ denote the production rate of all units in the system. The total production of the system is restricted to $\pi$, i.e., $ \sum_{i=1}^{n}u_i = \pi$. In practice this is relevant, since the production rate of a system could be set by the controlling party to, for, example, comply with a contracted output rate. Adding this restriction results in the following action space for the production
$$
U = \left\{ u \mid  \sum_{i=1}^{n}u_i = \pi \right\},
$$
Note that the production action space decreases from $p^n$ elements to ${n + \pi - 1} \choose{\pi - 1} $ by restriction the total production rate to the value $\pi$.  The entire action space consisting of the maintenance and production action space is given by
$$
A = R \times U 
$$

To define the immediate rewards and transition probabilities for the system, we define a function
$$
c(x_i,r_i) = \begin{cases}
c_{\text{pm}} & \text{ if } r_i = \text{M} \text{ and } x_i \neq L, \\
c_{\text{cm}} & \text{ if } r_i = \text{M} \text{ and } x_i = L, \\
0 & \text{ elsewhere}.
\end{cases}
$$
for all states $x_i$ and maintenance actions $r_i$ of unit $i \in \mathcal{I}$.  The state and action dependent immediate cost $\bar{c}(x,r)$ for the system is given by
$$
\bar{c}(x,r) = \begin{cases}
C + \sum_{i=1}^{n}c(x_i, r_i), &\text{if } r \text{ contains a maintenance action},\\
0, &\text{otherwise}.
\end{cases}
$$
Note that if we select action $a = (u, r)$ in state $x$, then $\bar{c}(x, a) = \bar{c}(x,r)$.

\subsection{Deterioration Process}
The transition probabilities for all units are identical and independent. Each unit $i$ deteriorates according to a Poisson process with rate $\lambda(u_i)$, which depends on the current production rate of the unit. This function for a given production rate $u_i$ is given by
$$
\lambda(u_i) = g(u_i) \lambda_{\text{max}},
$$
where $\lambda_{\text{max}}$ is the rate parameter under the maximum production rate, and $g$ is some non-decreasing function. For now we assume that $g(u_i) = u_i$. For a unit $i$ the probability of having $q \geq 0$ jumps in the deterioration states for a time epoch under production rate $u_i$ is given by
$$
p(u_i,q) = \frac{\lambda(u_i)^q}{q!}e^{-\lambda(u_i)}.
$$
The transition probability of unit $k$ for going from state $i$ to $j$ with production rate $u_i$ and maintenance action DN for any time epoch is defined as
$$
P_{u_k,\text{DN}}(i,j) = \begin{cases}
p(u_k, j-i) &\text{ if } i \leq j \leq L-1, \\
0 &\text{ if } j < i, \\
1 - \sum_{l=0}^{L-1-i}p(u_k,l) & \text{ if } i \leq m-1  \text{ and } j = L, \\
1 & \text{ if } i = j = L,
\end{cases}
$$
and for maintenance action M by
$$
P_{u_k,\text{M}}(i,j) = \begin{cases}
	p(u_k, j) &\text{ if } j \leq  L-1, \\
	1 - \sum_{l=0}^{L-1}p(u_k,l) & \text{ if } j = L.
\end{cases}
$$
Note that for the maintenance action M the probability of going from state $i$ to $j$ does not depend on $i$, since we assume that the system will go to state 0 instantaneously.

The transition probability of a unit in the system is independent of the other units. Therefore, the transition probability that the system goes from state $x = (x_1, \dots, x_n)$ to state $\tilde{x} = (\tilde{x}_1, \dots, \tilde{x}_n)$ by selecting production rates $u$ and maintenance actions $m$ is given by
$$
p(x, \tilde{x} \mid u, m) = \prod_{i = 1}^nP_{u_i,m_i}(x_i, \tilde{x}_i),
$$
due to independence of the deterioration process of the units.

Let $\boldsymbol{\sigma} = (\sigma, \sigma, \dots)$ be a stationary deterministic policy. This policy induces the cost process $\{X_t, \bar{c}(X_t, d(X_t))\}$, where $d(X_t)$ is the action for state $X_t$ corresponding to policy $\boldsymbol{\sigma}$. The average cost $g$ for this problem with policy $\boldsymbol{\sigma}$ for initial value $X(0) = x$ is given by
$$
g^{\boldsymbol{\sigma}}(x) = \lim_{N \to \infty}\frac{1}{N}\left \{\mathbb{E}^{\boldsymbol{\sigma}}_x \sum_{i=0}^N\bar{c}(X_i, d(X_t)) \right\}.
$$
Our goal is to minimise to find a policy $\boldsymbol{\sigma}^*$ such that the average cost $g$ is minimised. A Successive Approximation algorithm can be used to find the minimising policy.

\section{Extensions}
A list of extensions (ordered by priority) is given below:
\begin{enumerate}
	\item Let the deterioration process of the units be modelled by a gamma process.
	\item Consider other forms for the function $g$, e.g., see what happens when this function is convex or concave.
	\item Study a system with more than two units.
	\item Add revenue instead of imposing a restriction on the total production.
	\item Add lead time for maintenance.
	\item Study dependency of deterioration process of units. For windmills one could imagine that the deterioration is partially determined by weather conditions which are identical for windmills within a wind park.
	\item Consider imperfect condition information.
\end{enumerate}

\end{document}

### 2018-05-11 src/init/g.cpp
First attempt for Init class

@@ -1,0 +1,10 @@
+#include "init.ih"
+
+/*
+ * Define the function g
+ */
+
+num_t Init::g(double prodRate) const
+{
+  return d_beta + (1 - d_beta) * pow(prodRate, d_alpha);
+}

### 2018-05-11 src/init/init.h

@@ -1,0 +1,55 @@
+#ifndef INCLUDED_INIT_
+#define INCLUDED_INIT_
+
+#include <vector>
+
+typedef double num_t;
+
+class Init
+{
+    // Maintenance costs
+    num_t d_ccm;
+    num_t d_cpm;
+    num_t d_C;
+
+    // State space
+    size_t  d_nStates;
+    num_t   d_dL;
+
+    // Action space
+    num_t               d_pi;
+    size_t              d_nRates;
+    std::vector<num_t>  d_prodRates;
+
+    // Probabilities
+    num_t d_mu;
+    num_t d_var;
+    num_t d_dT;
+
+    std::vector<std::vector<num_t>> d_prop1;
+    std::vector<std::vector<num_t>> d_prop2;
+
+    num_t d_beta;
+    num_t d_alpha;
+    
+    std::vector<num_t> d_detRates;
+
+    // Costs
+    std::vector<num_t> d_terminalCosts;
+    
+    
+    public:
+        Init();
+
+        void    initialise();
+
+    private:
+        num_t   g(double prodRate) const;
+        void    printTransitionMatrix() const;
+        void    setTransitionGamma();
+        void    setCost();
+        void    setActions();
+        void    setActions(num_t prodRates);
+};
+
+#endif 

### 2018-05-11 src/init/printtransitionmatrix.cpp

@@ -1,0 +1,29 @@
+#include "init.ih"
+
+/* 
+ * Print the transition matrix for all actions
+ */
+
+void Init::printTransitionMatrix() const
+{
+  if (d_prop1.size() == 0 or d_prop2.size() == 0)
+    throw string("Error: transition matrices are unset (Init::printTransitionMatrix.cc)");
+  
+  cout << setprecision(3) << fixed;
+  for (size_t adx = 0; adx < d_nRates; ++adx)
+  {
+    auto const &p1 = d_prop1[adx];
+    auto const &p2 = d_prop2[adx];
+    
+    cout << " a = " << adx << endl;
+    
+    for (size_t idx = 0; idx < d_nStates; ++idx)
+        cout << setw(7) << p1[idx];
+    
+    cout << " - " << accumulate(begin(p1), end(p1), 0.0) << endl;
+    
+    for (size_t idx = 0; idx < d_nStates; ++idx)
+        cout << setw(7) << p2[idx];
+    cout << endl << endl;
+  }
+}

### 2018-05-11 src/init/setactions.cpp

@@ -1,0 +1,28 @@
+#include "init.ih"
+
+/* 
+ * First the possible production rates are set
+ * These are 'uniformly' distributed over the interval [0,1]
+ * 
+ * Second, the corresponding relative deterioration rates are set
+ * 
+ */
+
+void Init::setActions()
+{
+  d_prodRates = bs::seq_length(0.0, 1.0, d_nRates);
+  d_detRates  = vector<num_t>(d_nRates, 0);
+  
+  for (size_t idx = 0; idx < d_nRates; ++idx)
+    d_detRates[idx] = g(d_prodRates[idx]); 
+}
+
+/*
+ *   Alternatively, we only have full or no production
+ */
+
+void Init::setActions(num_t prodRate)
+{
+  d_prodRates = {{0.0, prodRate}};
+  d_detRates  = {{g(0), g(prodRate)}};
+}

### 2018-05-11 src/init/setcosts.cpp

@@ -1,0 +1,13 @@
+#include "init.ih"
+
+/* 
+ * This function set the terminal cost vector
+ * i.e., the costs for being in a specific state at the end of the time horizon
+ * The datamember d_terminalCosts is given to the MDP class
+ */
+
+void Init::setCost()
+{
+    d_terminalCosts = vector<num_t>(d_nStates, d_cpm);
+    d_terminalCosts.back() = d_ccm;
+}

### 2018-05-11 src/init/settransitiongamma.cpp

@@ -1,0 +1,62 @@
+#include "init.ih"
+
+#include <boost/math/distributions/gamma.hpp>
+
+/* 
+ * Set the member that contains the transition matrix
+ * For each action there are two transition vectors of size nStates
+ * The first one gives the probability of 'idx' jumps
+ * The second one gives the probability of an instant failure for state 'idx'
+ * 
+ * 
+ * TODO
+ * correct for time length
+ */
+
+
+void Init::setTransitionGamma()
+{
+  // Create empty transition vectors
+  // d_prop1 gives the probability of a jump with size idx
+  // d_prop2 gives the probability of an instant failure from state idx
+  
+  d_prop1 = vector<vector<num_t>> (d_nRates, vector<num_t> (d_nStates, 0));
+  d_prop2 = vector<vector<num_t>> (d_nRates, vector<num_t> (d_nStates, 0));
+  
+  // Fill the transion vectors for each action (= production rate)
+  for (size_t adx = 0; adx < d_nRates; ++adx)
+  {
+    num_t const shape = d_dT * (d_mu * d_mu) / d_var;
+    num_t const scale = d_detRates[adx] * (d_var / d_mu);
+    
+    // Find jump probabilities
+    if (d_mu * d_detRates[adx] == 0)
+      d_prop1[adx][0] = 1;
+
+    else
+    {
+      boost::math::gamma_distribution<num_t> distribution(shape, scale);
+      d_prop1[adx][0] = cdf(distribution, 0.5 * d_dL);
+      
+      for (size_t idx = 1; idx < d_nStates; ++idx)
+        d_prop1[adx][idx] = cdf(distribution, (idx + 0.5) * d_dL) - cdf(distribution, (idx - 0.5) * d_dL);
+    }
+    
+    // Probability of failure
+    for (size_t idx = 0; idx < d_nStates; ++idx)
+      d_prop2[adx][idx] = 1 - accumulate(begin(d_prop1[adx]), end(d_prop1[adx]) - idx - 1, 0.0);
+  }
+  
+  
+  
+  
+//   ofstream file("./jumps");
+//   file << fixed << setprecision(15);
+//   
+//   file << "0, " << d_nRates - 1 << '\n';
+//   
+//   for (size_t idx = 0; idx < d_nStates; ++idx)
+//     file << d_prop1[0][idx] << ", " << d_prop1[d_nRates - 1][idx] << '\n';
+//   
+//   throw;
+}

### 2018-05-11 src/mdp/converged.cpp
First attempt for MDP class

@@ -1,0 +1,11 @@
+#include "mdp.ih"
+
+bool MDP::converged()
+{
+    num_t span = bs::span(d_valueB, d_valueA);
+
+    if (span < d_eps)
+        return true;
+    
+    return false;
+}

### 2018-05-11 src/mdp/exp_cost.cpp

@@ -1,0 +1,13 @@
+#include "mdp.ih"
+
+/*
+    For every state we calculate the new value 
+    to obtain V_t+1
+*/
+
+void MDP::exp_cost()
+{
+    for (size_t x1 = 0; x1 != d_state_size; ++x1)
+        for (size_t x2 = 0; x2 != d_nStates; ++x2)
+            d_valueB[hash(x1, x2)] = expected_cost(x1, x2);
+}

### 2018-05-11 src/mdp/expected_cost.cpp

@@ -1,0 +1,13 @@
+#include "mdp.ih"
+
+/*
+    Calculate the expected cost for a given state 
+*/
+
+size_t MDP::expected_cost(size_t state1, size_t state2)
+{
+    num_t tmpVal = 0;
+    num_t bestVal = 0;
+
+
+}

### 2018-05-11 src/mdp/expected_costs.cpp

@@ -1,0 +1,13 @@
+#include "mdp.ih"
+
+/*
+ *  Calculate the expected cost for a given state and action
+ */
+
+size_t MDP::expected_costs(size_t state1, size_t state2, num_t prodRate1, prodRate2)
+{
+    num_t tmpVal = 0;
+    num_t bestVal = 0;
+
+
+}

### 2018-05-11 src/mdp/mdp.h

@@ -1,0 +1,35 @@
+#ifndef INCLUDED_MDP_
+#define INCLUDED_MDP_
+
+#include <vector>
+
+#include "../init/init.h"
+
+class MDP
+{
+    std::vector<num_t> d_valueA;
+    std::vector<num_t> d_valueB;
+
+    size_t  d_nStates;
+    num_t   d_eps;
+
+    public:
+        MDP(Init const &init);
+
+        void solve();
+
+    private:
+        bool    converged();
+        void    exp_cost();
+        num_t   expected_cost(size_t state1, size_t state2);
+        num_t   expected_costs(size_t state1, size_t state2, num_t prodRate1, num_t prodRate2);
+
+        size_t hash(size_t state1, size_t state2);
+};
+
+inline size_t MDP::hash(size_t state1, size_t state2) 
+{
+    return state1 * d_nStates + state2;
+}
+
+#endif

### 2018-05-11 src/mdp/solve.cpp

@@ -1,0 +1,16 @@
+#include "mdp.ih"
+
+/*
+ * Value iteration algorithm.
+ * Select v_0. 
+ * For each s in S, compute v_n+1.
+ */
+
+void MDP::solve()
+{
+    do 
+    {
+        // Compute for each s in S the optimal maintenance action.
+        
+    } (converged())
+}

### 2018-05-11 src/bs.h
Added header file containing useful (mathematical) functions

@@ -1,0 +1,52 @@
+#ifndef INCLUDED_BS_
+#define INCLUDED_BS_
+
+#include <vector>
+#include <algorithm>
+
+/*
+ * This file contains useful functions that will be used
+ * throughout the project
+ */
+
+namespace bs
+{
+    /*
+     * Function that will return a vector of size 'size' from 'begin' to 'end'.
+     */
+    std::vector<double> seq_length(double const begin, double const end, int const size)
+    {
+        std::vector<double> seq(size, 0);
+
+        seq[0] = begin;
+        for (size_t idx = 1; idx != (size-1); ++idx)
+            seq[idx] = begin + idx * ((end - begin) / size);
+        seq[size-1] = end;
+        
+        return seq;
+    }
+
+    /*
+     * Calculate the span of two vectors
+     * 
+     * TODO check if O(n) is fastest for obtaining differnence vector... I think so
+     */ 
+    double span(vector<double> const &first, vector<double> const &second)
+    {  
+        size_t length = first.size()
+        
+        if (length != second.size())
+            throw string("Error: vectors do not have same length (bs::span(first, second))");
+
+        std::vector<double> diff(length, 0);
+        for (size_t idx = 0; idx != length; ++idx)
+            diff[idx] = first[idx] - second[idx];
+        
+        auto result = std::minmax_element(diff.begin(), diff.end());
+
+        return result.second - result.first;
+    }
+}
+
+
+#endif

### 2018-05-11 src/bs/bs.cpp
Moved bs functions and fixed some bugs

@@ -1,22 +1,15 @@
-#ifndef INCLUDED_BS_
-#define INCLUDED_BS_
+#include "bs.h"
 
-#include <vector>
-#include <algorithm>
-
-/*
- * This file contains useful functions that will be used
- * throughout the project
- */
+using namespace std;
 
 namespace bs
 {
     /*
      * Function that will return a vector of size 'size' from 'begin' to 'end'.
      */
-    std::vector<double> seq_length(double const begin, double const end, int const size)
+    vector<double> seq_length(double const begin, double const end, int const size)
     {
-        std::vector<double> seq(size, 0);
+        vector<double> seq(size, 0);
 
         seq[0] = begin;
         for (size_t idx = 1; idx != (size-1); ++idx)
@@ -33,20 +26,17 @@ namespace bs
      */ 
     double span(vector<double> const &first, vector<double> const &second)
     {  
-        size_t length = first.size()
+        size_t length = first.size();
         
         if (length != second.size())
             throw string("Error: vectors do not have same length (bs::span(first, second))");
 
-        std::vector<double> diff(length, 0);
+        vector<double> diff(length, 0);
         for (size_t idx = 0; idx != length; ++idx)
             diff[idx] = first[idx] - second[idx];
         
-        auto result = std::minmax_element(diff.begin(), diff.end());
+        auto result = minmax_element(diff.begin(), diff.end());
 
         return result.second - result.first;
     }
 }
-
-
-#endif

### 2018-05-11 src/bs/bs.h

@@ -1,0 +1,19 @@
+#ifndef INCLUDED_BS_
+#define INCLUDED_BS_
+
+#include <vector>
+#include <algorithm>
+#include <iostream>
+
+/*
+ * This file contains useful functions that will be used
+ * throughout the project
+ */
+
+namespace bs
+{
+    std::vector<double> seq_length(double const begin, double const end, int const size);
+    double span(std::vector<double> const &first, std::vector<double> const &second);
+}
+
+#endif

### 2018-05-11 src/mdp/exp_cost.cpp

@@ -7,7 +7,7 @@
 
 void MDP::exp_cost()
 {
-    for (size_t x1 = 0; x1 != d_state_size; ++x1)
+    for (size_t x1 = 0; x1 != d_nStates; ++x1)
         for (size_t x2 = 0; x2 != d_nStates; ++x2)
             d_valueB[hash(x1, x2)] = expected_cost(x1, x2);
 }

### 2018-05-11 src/mdp/expected_cost.cpp

@@ -4,7 +4,7 @@
     Calculate the expected cost for a given state 
 */
 
-size_t MDP::expected_cost(size_t state1, size_t state2)
+num_t MDP::expected_cost(size_t state1, size_t state2)
 {
     num_t tmpVal = 0;
     num_t bestVal = 0;

### 2018-05-11 src/mdp/expected_costs.cpp

@@ -4,7 +4,7 @@
  *  Calculate the expected cost for a given state and action
  */
 
-size_t MDP::expected_costs(size_t state1, size_t state2, num_t prodRate1, prodRate2)
+num_t MDP::expected_costs(size_t state1, size_t state2, num_t prodRate1, num_t prodRate2)
 {
     num_t tmpVal = 0;
     num_t bestVal = 0;

### 2018-05-11 src/mdp/solve.cpp

@@ -12,5 +12,5 @@ void MDP::solve()
     {
         // Compute for each s in S the optimal maintenance action.
         
-    } (converged())
+    } while(converged());
 }

### 2018-05-11 src/run.cpp

@@ -1,0 +1,6 @@
+#include <iostream>
+
+int main(int argc, char**argv) 
+{
+    std::cout << "Hello world!\n";
+}

### 2018-05-12 src/init/init.cpp
Init constructor and initialise function created

@@ -1,0 +1,22 @@
+#include "init.ih"
+
+// Empty initialiser
+Init::Init()
+{}
+
+Init::Init(num_t ccm, num_t cpm, num_t C, size_t nStates, num_t dL, size_t pi,
+           size_t nRates, num_t mu, num_t var, num_t dT, num_t beta, num_t alpha)
+:
+    d_ccm(ccm),
+    d_cpm(cpm),
+    d_C(C),
+    d_nStates(nStates),
+    d_dL(dL),
+    d_pi(pi),
+    d_nRates(nRates),
+    d_mu(mu),
+    d_var(var),
+    d_dT(dT),
+    d_beta(beta),
+    d_alpha(alpha)
+{}

### 2018-05-12 src/init/init.h

@@ -17,7 +17,7 @@ class Init
     num_t   d_dL;
 
     // Action space
-    num_t               d_pi;
+    size_t              d_pi;
     size_t              d_nRates;
     std::vector<num_t>  d_prodRates;
 
@@ -40,6 +40,8 @@ class Init
     
     public:
         Init();
+        Init(num_t ccm, num_t cpm, num_t C, size_t nStates, num_t dL, size_t pi,
+             size_t nRates, num_t mu, num_t var, num_t dT, num_t beta, num_t alpha);
 
         void    initialise();
 

### 2018-05-12 src/init/initialise.cpp

@@ -1,0 +1,11 @@
+#include "init.ih"
+
+/*
+ * This functions sets all data members to final states
+ */
+void Init::initialise()
+{
+    setActions();
+    setCost();
+    setTransitionGamma();
+}

### 2018-05-12 src/mdp/exp_cost.cpp
Functions to calculate expected costs created

@@ -7,7 +7,7 @@
 
 void MDP::exp_cost()
 {
-    for (size_t x1 = 0; x1 != d_nStates; ++x1)
-        for (size_t x2 = 0; x2 != d_nStates; ++x2)
-            d_valueB[hash(x1, x2)] = expected_cost(x1, x2);
+    for (size_t state1 = 0; state1 != d_nStates; ++state1)
+        for (size_t state2 = 0; state2 != d_nStates; ++state2)
+            d_expectedCost[hash(state1, state2)] = expected_cost(state1, state2);
 }

### 2018-05-12 src/mdp/expected_cost.cpp

@@ -2,12 +2,23 @@
 
 /*
     Calculate the expected cost for a given state 
+
+    TODO obvious improvement is to restrict 'pdx2' to 'd_pi - pdx1'
+    We assume that '2*d_nRates > d_pi > d_nRates'
 */
 
 num_t MDP::expected_cost(size_t state1, size_t state2)
 {
     num_t tmpVal = 0;
-    num_t bestVal = 0;
+    num_t bestVal = numeric_limits<num_t>::max();
 
+    // Select minimum costs for all possible production states
+    for (size_t pdx1 = (d_pi - d_nRates - 1); pdx1 != d_nRates; ++pdx1)
+    {
+        tmpVal = expected_costs(state1, state2, pdx1, d_pi - pdx1);
+        if (tmpVal < bestVal)
+            bestVal = tmpVal;
+    }
 
+    return bestVal;
 }

### 2018-05-12 src/mdp/expected_costs.cpp

@@ -1,13 +1,33 @@
 #include "mdp.ih"
 
 /*
- *  Calculate the expected cost for a given state and action
+ *  Calculate the expected cost for a given state and action.
+ *  Loop over all possible jumps to states
  */
 
-num_t MDP::expected_costs(size_t state1, size_t state2, num_t prodRate1, num_t prodRate2)
+num_t MDP::expected_costs(size_t state1, size_t state2, size_t prodRate1, size_t prodRate2)
 {
-    num_t tmpVal = 0;
-    num_t bestVal = 0;
+    num_t costs = 0;
+    num_t prop1;
 
+    // Loop over all jump probabilities from 'state1' to 'd_nStates' and obtain costs.
+    for (size_t sdx1 = 0; sdx1 != (d_nStates-state1-1); ++sdx1) 
+    {
+        prop1 = d_prop1[prodRate1][sdx1];
+        for (size_t sdx2 = 0; sdx2 != (d_nStates-state2-1); ++sdx2)
+        {
+            costs += prop1 * d_prop1[prodRate2][sdx2] * d_valueB[hash(sdx1, sdx2)];
+        }
+        costs += prop1 * d_prop2[prodRate2][state2] * d_valueB[hash(sdx1, d_nStates)];
+    }
 
+    // Add costs when unit 1 goes to failed state;
+    prop1 = d_prop2[prodRate1][state1];
+    for (size_t sdx2 = 0; sdx2 != (d_nStates-state2-1); ++sdx2)
+    {
+        costs += prop1 * d_prop1[prodRate2][sdx2] * d_valueB[hash(d_nStates, sdx2)];
+    }
+    costs += prop1 * d_prop2[prodRate2][state2] * d_valueB[hash(d_nStates, d_nStates)];
+
+    return costs;
 }

### 2018-05-12 src/mdp/mdp.h

@@ -7,11 +7,32 @@
 
 class MDP
 {
+    // Values
     std::vector<num_t> d_valueA;
     std::vector<num_t> d_valueB;
 
-    size_t  d_nStates;
-    num_t   d_eps;
+    // Expected costs for every production rate
+    std::vector<num_t> d_expectedCost;
+
+    // Maintenance costs
+    num_t const d_ccm;
+    num_t const d_cpm;
+    num_t const d_C;
+
+    // State and action space
+    size_t const d_nStates;
+    size_t const d_pi;
+    size_t const d_nRates;
+
+    // Jump probabilities
+    std::vector<std::vector<num_t>> const d_prop1;
+    std::vector<std::vector<num_t>> const d_prop2;
+
+    // Costs
+    std::vector<num_t> const d_terminalCosts;
+
+    // Value iteration parameter
+    num_t const d_eps;
 
     public:
         MDP(Init const &init);
@@ -22,7 +43,7 @@ class MDP
         bool    converged();
         void    exp_cost();
         num_t   expected_cost(size_t state1, size_t state2);
-        num_t   expected_costs(size_t state1, size_t state2, num_t prodRate1, num_t prodRate2);
+        num_t   expected_costs(size_t state1, size_t state2, size_t prodRate1, size_t prodRate2);
 
         size_t hash(size_t state1, size_t state2);
 };

### 2018-05-12 src/mdp/solve.cpp
First version value iteration algorithm

@@ -10,7 +10,14 @@ void MDP::solve()
 {
     do 
     {
-        // Compute for each s in S the optimal maintenance action.
+        // Set new values to old values
+        d_valueA = d_valueB;
+        
+        // Compute the expected cost for every state and optimal production
+        exp_cost();
+
+        // Obtain new value 
+        value_cost();
         
     } while(converged());
 }

### 2018-05-12 src/mdp/value_cost.cpp

@@ -1,0 +1,8 @@
+#include "mdp.ih"
+
+void MDP::value_cost()
+{
+    for (size_t state1 = 0; state1 != d_nStates; ++state1)
+        for (size_t state2 = 0; state2 != d_nStates; ++state2)
+            d_valueB[hash(state1, state2)] = value_costs(state1, state2);
+}

### 2018-05-28 report/paper_scm.tex
Update report

\documentclass[a4paper,12pt]{article}

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage[]{geometry}
\usepackage{inconsolata}
\usepackage{courier}
\usepackage{listings}
\usepackage{enumerate}
\usepackage{float}
\usepackage{answers}
\usepackage[english]{babel}
\usepackage{commath}
\usepackage[labelfont=bf]{caption}
\usepackage{mwe}
\usepackage[longnamesfirst]{natbib}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{eurosym}
\usepackage{siunitx}
\usepackage{ctable}
\usepackage{subfig}
\usepackage{dsfont}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}

\title{Maintenance Policies for a Multi-Unit System with Adjustable Production Rates}
\author{Mark van der Broek}
\date{\today}

\begin{document}

	\maketitle

\section{Introduction}	
% Good maintenance policies are required

% For systems with multiple units, there exists dependence between the units.

% Windmills require expensive fixed costs, so clustering maintenance can be profitable. In many windmill contracts a specified production has to be met for the total windpark. Try to combine production with maintenance policies.

The system also allows to adjust the production rate of each unit, which has an impact on the deterioration process of the unit. The units deteriorate independently of each other, and we can perfectly observe this deterioration level for each unit at any moment in time. An graphical overview of the order of events within a time epoch is presented in Figure 1.

For the analysis I will consider a system of two units, and if time permits we will relax this assumption to consider a system with more units. For the maintenance policy can perform maintenance or not with cost $c_\text{pm}$ if the unit has not failed yet and cost $c_\text{cm}$ if it has failed. The cost of preventive maintenance is assumed to be less than corrective maintenance cost ($c_\text{pm} < c_\text{cm}$). Furthermore, for any time period we do perform maintenance a fixed cost $C$ has to be incurred. Therefore, depending on the size of these three costs it could be beneficial to cluster the maintenance actions.

\section{Markov Decision Process}
In this section we provide the Markov Decision Process (MDP) formulation of our problem. At the start of a time unit, the system can be in a certain state. Based on this current state, a set of maintenance actions can be performed. For each unit, this set consists of doing nothing or performing maintenance. After selecting the maintenance action, we set the production rate of each unit. Based on the current state, maintenance and production action for a time period, each unit of the system deteriorates according to some transition probabilities. A schematic overview of these processes within a time unit are presented in Figure \ref{timeline}. In this section we will specify the state space, action space, transition probabilities and the expected costs.

\begin{figure}[H]
	\centering
	\begin{tikzpicture}
	\coordinate (t) at (0,0);
	\coordinate (X) at (1,0);
	\coordinate (R) at (2,0);
	\coordinate (M) at (3.5,0);
	\coordinate (U) at (5,0);
	\coordinate (UE) at (9,0);
	\coordinate (D) at (10,0);
	\coordinate (t+1) at (11,0);

	% Draw the time horizon
	\draw[thick, |-|] (t) -- (t+1);
	\draw (t) node[below=3pt] {$t$};
	\draw[thick, ->] (1,1) node[above, align=center] {Observe $X_t$} -- (X);
	\draw[thick, ->] (2,-1) node[below, align=center] {Maintenance \\ decision $r$} -- (R);
	\draw[thick, ->] (3.5, 1) node[above, align=center] {Perform \\ maintenance} -- (M);
	\draw[thick, ->] (5, -1) node[below, align=center] {Production \\ decision $u$} -- (U);
	\draw[thick, decorate, decoration={brace,raise=2pt, amplitude=3pt}] (6,0) -- (9, 0) node[sloped,midway, above=6pt] {Produce};
	\draw[thick, ->] (10, 1) node[above, align=center] {System deteriorates} -- (D);
	\draw (t+1) node[below=3pt] {$t+1$};
	\end{tikzpicture}
	\caption{Order of events for time epoch $t$.}
\end{figure}

\subsection{State space}
We consider a multi-unit system that consists of $n$ identical units. We denote the set of units as $\mathcal{I} = \{1, \dots, n\}$. Each unit deteriorates according to a discrete-time Markov chain with deterioration states $X = \{0, 1, \dots, L\}$. State $0$ represents the as-good-as-new state and state $L$ the failed state. The whole state of all units is denoted by
$$
x = (x_1, \dots, x_n) \in X^n.
$$

\subsection{Action space}
At the start of each time unit we have to decide on which components to maintain (M) and do nothing (DN). Maintenance can be performed one all units, independent of the current state, i.e., the action space for the maintenance decision can be defined as
$$
R = \left \{r = (r_1, \dots, r_n): r_i \in \{\text{DN}, \text{M}\}, \enskip \forall i \in \mathcal{I} \right \}.
$$

After selecting the maintenance action, we also have to set the production rate of each unit. Each unit can produce at most $m+1$ different production rates, ranging from $0$ to $1$ (maximum production). The possible production rates are uniformly distributed over the interval. If a unit is in the failed state, we cannot produce, so the production rate of this unit is set to $0$. For all other deteriorating levels, we can select one of the $m+1$ production rates. We define the set of possible production rates for unit $i$ in state $x_i$ $U(x_i)$ as follows
$$
U(x_i) = \begin{cases}
\{0, 1/m, \dots, 1\} & \text{ if } x_i \neq L, \\
\{0\} & \text{ if } x_i = L.
\end{cases}
$$
Hence, the 
Let $u = (u_1, \dots, u_n)$ denote the production rate of all units in the system. The total production of the system is restricted to $\pi$, i.e., $ \sum_{i=1}^{n}u_i = \pi$. This results in the following action space for the system in state $x$
$$
U(x) = \left\{ u  :  \sum_{i=1}^{n}u_i = \pi, u_i \in U(x_i) \enskip \forall i \in \mathcal{I} \right\}.
$$
Note that the production action space decreases from $(m+1)^n$ elements to ${n + \pi - 1} \choose{\pi - 1} $ by restriction the total production rate to the value $\pi$.

\subsection{Deterioration Process}
The transition probabilities for all units are identical and independent. Each unit $i$ deteriorates according to a gamma process with rate $\lambda(u_i)$, which depends on the current production rate of the unit, and variance $\sigma^2$. The rate for a given production rate $u_i$ is given by
$$
\lambda(u_i) = g(u_i) \lambda_{\text{max}},
$$
where $\lambda_{\text{max}}$ is the rate parameter under the maximum production rate, and $g$ is some non-decreasing function. We assume that $g(u_i)$ as a function of the production rate $u_i$ for unit $i$ is of the following form
$$
g(u_i) = \beta + (1-\beta)u_i^\alpha.
$$
We assume that $0 \leq \beta \leq 1$ and $\alpha \geq 0$, since $g$ should be non-decreasing in $u_i$. \textbf{[[Add some information about the shape of $g$]]}

For a unit $i$ the probability of having $q \geq 0$ jumps in the deterioration states for a time epoch under production rate $u_i$ is given by $p(u_i,q)$, which can be found by discretising the gamma process (see De Jonge). The transition probability of unit $k$ for going from state $i$ to $j$ with production rate $u_i$ and maintenance action DN for any time epoch is defined as
$$
P_{u_k,\text{DN}}(i,j) = \begin{cases}
p(u_k, j-i) &\text{ if } i \leq j \leq L-1, \\
0 &\text{ if } j < i, \\
1 - \sum_{l=0}^{L-1-i}p(u_k,l) & \text{ if } i \leq m-1  \text{ and } j = L, \\
1 & \text{ if } i = j = L,
\end{cases}
$$
and for maintenance action M by
$$
P_{u_k,\text{M}}(i,j) = \begin{cases}
p(u_k, j) &\text{ if } j \leq  L-1, \\
1 - \sum_{l=0}^{L-1}p(u_k,l) & \text{ if } j = L.
\end{cases}
$$
Note that for the maintenance action M the probability of going from state $i$ to $j$ does not depend on $i$, since we assume that the system will go to state 0 immediately after the maintenance action.

The transition probability of a unit in the system is independent of the other units. Therefore, the transition probability that the system goes from state $x = (x_1, \dots, x_n)$ to state $\tilde{x} = (\tilde{x}_1, \dots, \tilde{x}_n)$ by selecting production rates $u$ and maintenance actions $m$ is given by
$$
p(x, \tilde{x} \mid u, m) = \prod_{i = 1}^nP_{u_i,m_i}(x_i, \tilde{x}_i),
$$
due to independence of the deterioration process of the units.

\subsection{Expected costs}
To define the cost for a time unit, we define a function
$$
c(x_i,r_i) = \begin{cases}
c_{\text{pm}} & \text{ if } r_i = \text{M} \text{ and } x_i \neq L, \\
c_{\text{cm}} & \text{ if } r_i = \text{M} \text{ and } x_i = L, \\
0 & \text{ elsewhere}.
\end{cases}
$$
for all states $x_i \in X$ and maintenance actions $r_i \in \{\text{DN}, \text{M}\}$ of unit $i \in \mathcal{I}$. This means that if unit $i$ is in the failed state and we decide to perform maintenance, we incur the corrective maintenance costs $c_{\text{cm}}$, and if the unit has not failed, then the preventive maintenance costs $c_{\text{pm}}$ is incurred.

The expected costs for a time unit by selecting maintenance action $r$ and in state $x$ is given by
$$
\bar{c}(x,r) = \begin{cases}
C + \sum_{i=1}^{n}c(x_i, r_i), &\text{if } r \text{ contains a maintenance action},\\
0, &\text{otherwise}.
\end{cases}
$$
In words, this means that if we do not perform maintenance on any unit for a given time period, we do not incur any costs. However, if we do perform maintenance, we incur the fixed cost $C$ and some corrective or preventive costs depending on the state of the maintained unit.
\subsection{Performance criteria}
As a performance criterion, we are interested in minimizing the long-run average cost per time unit. In this way, we can quantify the impact of adding an additional component, and find the most cost-efﬁcient maintenance policy.

\section{Numerical results}

\subsection{Two unit setting}

\subsection{Adding additional unit}
\begin{enumerate}
	\item Directly compare the results for different mean rates $g$. 
\end{enumerate}

\subsection{Benchmark policies}
Since the action space of our problem consists of a maintenance action and a production action, we distinguish between a benchmark for the maintenance action and the production action.

In the literature, condition-based maintenance is often implemented in the form of a threshold policy (CITE). Due to the economic dependence (fixed cost of maintenance) and the liberty to set the production rate of the components, this policy is probably not optimal. For this policy, if the deterioration level of a unit is at or exceed a threshold level $T_r$, we perform a maintenance action. We find the optimal $T_r$ by simulating the system for a grid of values for $T_r$.

Production benchmark ideas:
\begin{enumerate}
	\item Share production equally between all units for all states.
	\item Random pool of $n$ units that produce equal amount.
	\item Best 50\% produces 80\% and worst 50\% 20\%.
\end{enumerate}

\subsection{Sensitivity analysis}
\begin{enumerate}
	\item Effect of fixed costs. More or less clustered maintenance.
	\item Effect of variance.

\end{enumerate}

\appendix

\section{Appendix}

\subsection{Discretising the gamma process}

### 2018-05-28 src/solve_experiments.h

@@ -10,7 +10,7 @@
 void solve_experiment(Init const &init)
 {
     // Initialise MDP
-    MDP mdp{init};
+    MDP mdp{init, false};
     mdp.solve();
     mdp.write_solution();
     std::cout << "Solved experiment " << mdp.get_id() << std::endl;
@@ -41,6 +41,5 @@ void solve_from_file()
     // Solve all experiments in different threads
     for (Init const &init: init_vec)
         solve_experiment(init);
-
 }
 #endif

### 2018-05-29 create_experiments.R
Start with policy evaluator

@@ -9,10 +9,10 @@ grid_size = 10
 base_experiment = tibble(ccm = 1.0,
                          cpm = 0.2,
                          C = 0.5,
-                         nStates = 50,
+                         nStates = 30,
                          dL = 1.0,
-                         pi = 60,
-                         nRates = 50,
+                         pi = 40,
+                         nRates = 30,
                          mu = 1.0,
                          var = 2.0,
                          dT = 1.0,
@@ -35,9 +35,14 @@ create_experiments = function(variableName, values) {
 }
 
 # Variance experiments
-var_options = seq(0.1, 5, length.out = grid_size)
+var_options = seq(0.5, 5, length.out=grid_size)
 experiments = experiments %>%
   bind_rows(create_experiments("var", var_options))
+
+# Fixed cost effect
+fixed_cost_options = seq(0, 3, length.out=grid_size)
+experiments = experiments %>%
+  bind_rows(create_experiments("C", fixed_cost_options))
 
 
 #### Output the experiments table ####

### 2018-05-29 create_figures.R

@@ -5,6 +5,13 @@
 library(tidyverse)
 library(reshape2)
 
+source("create_experiments.R")
+experiments = mutate(experiments, id = seq(0, nrow(experiments)-1))
+
+experiment_dir = "output"
+
+#### Create optimal policy plots ####
+
 # Function to map integer maintenance to string
 maintenance_int_to_string = function(maintenance_int) {
   if (maintenance_int == 0) {
@@ -16,39 +23,93 @@ maintenance_int_to_string = function(maintenance_int) {
 maintenance_int_to_string = Vectorize(maintenance_int_to_string, "maintenance_int")
 
 # Load optimal policy
-optimalPolicy = read_delim("output/optimal_policy.csv", delim=" ") %>%
-  mutate(`Maintenance action` = paste(maintenance_int_to_string(maintenance1), 
-                                 maintenance_int_to_string(maintenance2), 
-                                 sep=" ")) %>%
-  mutate(`Production action` = paste(production1, production2, sep=" ")) %>%
-  mutate(state1_max = state1 + 1,
-         state2_max = state2 + 1)
+experiment_names = list.files(experiment_dir, pattern = "^optimal_policy[0-9]*.csv$")
+n_experiments = length(experiment_names)
+idx = 0
 
-# In wide format
-optimalPolicy_wide = dcast(optimalPolicy, state1 ~ state2, value.var="Maintenance action")
+optimal_policy_data = NULL
+for (experiment_name in experiment_names) {
+  optimal_policy = read_delim(paste(experiment_dir, experiment_name, sep="/"), delim=" ") %>%
+    mutate(`Maintenance action` = paste(maintenance_int_to_string(maintenance1), 
+                                        maintenance_int_to_string(maintenance2), 
+                                        sep=" ")) %>%
+    mutate(`Production action` = paste(production1, production2, sep=" ")) %>%
+    mutate(state1_max = state1 + 1,
+           state2_max = state2 + 1,
+           id = idx)
+  
+  optimal_policy_data = optimal_policy_data %>%
+    bind_rows(optimal_policy)
+  idx = idx + 1
+}
 
-# Optimal maintenance with rectangles
-opt_maintenance_plot2 = ggplot(optimalPolicy, aes(xmin=state1, xmax=state1_max,
-                                                 ymin=state2, ymax=state2_max,
-                                                 fill=`Maintenance action`)) +
[…]

### 2018-05-29 src/mdp/applied_policy.cpp

@@ -1,0 +1,32 @@
+#include "mdp.ih"
+
+/*
+ * Evaluate cost of policy using a simulation of length
+ * 'simTime'.
+ * 
+ * TODO do we let the deterioration state be continuous?
+ * Or do we determine the thresholds of our maintenancePolicy?
+ */
+
+num_t MDP::applied_policy(std::vector<std::array<int, 2>> maintenancePolicy,
+                          std::vector<std::array<int, 2>> productionPolicy,
+                          int simTime) const
+{
+    double totalCosts = 0;
+    double detState1 = 0;
+    double detState2 = 0;
+
+    // Run simulation for simTime time units
+    for (int t = 0; t != simTime; ++t)
+    {
+        // Check if we perform maintenance
+
+        // Set production rate
+
+        // Determine deterioration (DISCRETE OR CONTINUOUS)
+
+    }
+
+    // Mean cost per time unit is total cost divided by length of time
+    return (totalCosts / simTime);
+}

### 2018-05-29 src/mdp/mdp.h

@@ -54,9 +54,12 @@ class MDP
         MDP() = delete;
         MDP(Init const &init, bool jumpHeuristic = false);
 
-        void solve();
-        void write_solution() const;
-        void print_all_info() const;
+        void    solve();
+        void    write_solution() const;
+        void    print_all_info() const;
+        num_t   applied_policy(std::vector<std::array<int, 2>> maintenancePolicy,
+                               std::vector<std::array<int, 2>> productionPolicy,
+                               int simTime) const;
         
         size_t get_id() const;
 

### 2018-05-30 report/paper_scm.tex
Update report

@@
 % Good maintenance policies are required
+Maintenance cost can constitute a large part of the total production costs.
+For example, \cite{GRABER2004861} states that maintenance costs can be at much as 30 percent of production costs for power plants, while \cite{BEVILACQUA200071} give an estimate of between 15 and 70 percent of the total production costs.
+For this reason, maintenance strategies have been extensively researched in the past decades.
+Reviews of past research have been published (CITE).
+Condition-based maintenance (CBM) is an advanced maintenance strategy that is based on performance and/or parameter monitoring and basing subsequent actions on the monitoring.
 
+Condition-based maintenance policies have been proposed for multi-unit systems with economic dependence \citep{OLDEKEIZER2016531}.
+\cite{OLDEKEIZER2018319} extend this research by including load sharing between units, that is, the deterioration of the functioning units is affected by the number of functioning units, where it is assumed that the load is shared uniformly over all functioning units.
+In this paper we consider a system where we relax this assumption and allow the load of all functioning units to depend on the current deterioration level of these units.
+
+The coupling of production and maintenance has been the subject of several studies in recent years.
+Li and Ni (2009 Jin, X., L.
+Li, and J.
+Ni.
+2009.
+“Option Model for Joint Production and Preventive Maintenance System.” International Journal of Production Economics 119(2): 347–353.
+[Crossref], [Web of Science ®], [Google Scholar]
+) present an option-based cost model for scheduling a joint production and maintenance policy for a single-period, single-product system.
+
+Offshore wind farms is one of the renewable energy solutions.
+Since the operation and maintenance costs represent a substantial portion of the total life cycle costs of these offshore wind farms \cite{hau2003wind}, reliability and condition strategies have drawn increasing interest for decreasing these costs, e.g.
+\cite{krokoszinski2003efficiency}.
+Existing CBM methods for offshore wind farm systems deal with wind turbines as individual components separately \citep{avner2007reliability}.
+Later maintenance strategies where economic dependence for maintenance actions between wind turbines is taken into account have gained interest \citep{tian2011condition}.
+The authors claim a maintenance costs reduction of up to 44.42\% compared to maintenance strategies that do not take the economic dependence into account.
+This reduction is mostly due to the fact that it exploits the large fixed costs from and cluster maintenance actions.
+One element of the fixed costs component is the cost of sending a maintenance team to the wind farm,, which generally requires expensive vessels.
+
+In this paper, we consider a joint production and maintenance strategy for a multi-unit system.
+We formulate the system as a Markov Decision Process and analyse the structural properties of the optimal policies.
+Our model extends the model of \cite{OLDEKEIZER2018319} by allowing the load sharing to be set for all units.
+The deterioration process of the units is modelled by a discrete-time Markov chain with a known transition probability matrix.
+The load sharing and maintenance strategy are jointly optimised, while in traditional methods the load sharing is set be uniform for all units.
+
+The remainder of the paper is organised as follows.
+Section 2 provides the definition and formulation of the system.
+Section 3 provides a description of the Markov Decision Process used to model the system introduced in Section 2.
+In Section 4 and 5 the structural properties of the optimal policies of a system with respectively 2 and 3 units is analysed.
+Conclusions and future research directions are provided in Section 6.
 % For systems with multiple units, there exists dependence between the units.
@@
 
-The system also allows to adjust the production rate of each unit, which has an impact on the deterioration process of the unit.
-The units deteriorate independently of each other, and we can perfectly observe this deterioration level for each unit at any moment in time.
-An graphical overview of the order of events within a time epoch is presented in Figure 1.
-
-For the analysis I will consider a system of two units, and if time permits we will relax this assumption to consider a system with more units.
-For the maintenance policy can perform maintenance or not with cost $c_\text{pm}$ if the unit has not failed yet and cost $c_\text{cm}$ if it has failed.
-The cost of preventive maintenance is assumed to be less than corrective maintenance cost ($c_\text{pm} < c_\text{cm}$).
-Furthermore, for any time period we do perform maintenance a fixed cost $C$ has to be incurred.
-Therefore, depending on the size of these three costs it could be beneficial to cluster the maintenance actions.
-
-\section{Markov Decision Process}
-In this section we provide the Markov Decision Process (MDP) formulation of our problem.
-At the start of a time unit, the system can be in a certain state.
-Based on this current state, a set of maintenance actions can be performed.
-For each unit, this set consists of doing nothing or performing maintenance.
-After selecting the maintenance action, we set the production rate of each unit.
-Based on the current state, maintenance and production action for a time period, each unit of the system deteriorates according to some transition probabilities.
-A schematic overview of these processes within a time unit are presented in Figure \ref{timeline}.
-In this section we will specify the state space, action space, transition probabilities and the expected costs.
+\section{System formulation}
+We consider a discrete-time system consisting of $N$ identical components, which are subject to load sharing and economic dependence.
+We denote the set of units as $\mathcal{I} = \{1, \dots, n\}$.
+The total load that has to be met during every time period is denoted $\pi$.
+Let $x_i$ denote the deterioration state of unit $i$, which deteriorates according to a discrete-time Markov chain with deterioration states $X = \{0, 1, \dots, L\}$.
+State $0$ represents the as-good-as-new state, $L-1$ denotes the most deteriorated state such that the unit still functions and state $L$ the failed state.
+We assume that the load capacity of a unit is not affected by the deterioration state.
+A schematic overview of the events in one time period is presented in Figure 1.
 
@@
 
+\subsection{Maintenance and load actions}
+The deterioration state of each unit is known at the beginning of each time period.
+This assumption becomes is justified for an increasing number of applications, due to better monitoring equipment.
+Maintenance is assumed to take negligible time and is performed after the deterioration states of each unit is observed.
+Preventive maintenance is performed at a cost of $c_\text{pm}$ if the unit has not reached the failed state, whereas a corrective maintenance action is performed in the failed state at a cost of $c_\text{cm}$.
+Generally, the cost of preventive maintenance is lower than the corrective maintenance action ($c_\text{pm} < c_\text{cm}$).
+Performing maintenance on a unit instantanously brings it back to the as-good-as new state.
+The economic dependence is included in the form of a fixed cost $C$ for performing maintenance for a time period.
+The cost is incurred when at least one unit will be maintained.
+
+After selecting the maintenance action, we also have to set the production rate of each unit.
+Each unit can produce at most $m+1$ different production rates, ranging from $0$ to $1$ (maximum production).
+The possible production rates are uniformly distributed over the interval.
+If a unit is in the failed state, we cannot produce, so the production rate of this unit is set to $0$.
+For all other deteriorating levels, we can select one of the $m+1$ production rates.
+
+\subsection{Deterioration Process}
+Each unit $i$ deteriorates according to a gamma process with rate $\lambda(u_i)$, which depends on the current production rate of the unit, and scale parameter $\gamma$.
+The rate for a given production rate $u_i$ is given by
+$$
+\lambda(u_i) = g(u_i) \lambda_{\text{max}},
+$$
+where $\lambda_{\text{max}}$ is the rate parameter under the maximum production rate, and $g$ is a non-decreasing function.
+We assume that $g(u_i)$ as a function of the production rate $u_i$ for unit $i$ is of the following form
+$$
+g(u_i) = \beta + (1-\beta)u_i^\alpha.
+$$
+We assume that $0 \leq \beta \leq 1$ and $\alpha \geq 0$, since $g$ should be non-decreasing in $u_i$.
+If the production rate $u_i$ is zero, then $g$ has value $\beta$ i.e., $\beta$ determines the deterioration rate if the unit does not produce.
+The parameter $\alpha$ determines the curvature of the function $g$.
+In Figure 2 the function $g$ is plotted for several settings of $\alpha$ and $\beta$.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.5\linewidth]{figures/g}
+	\caption{Function $g$ for values $\alpha = 0, 0.5, 1, 1.5$ and $\beta = 0.1$.}
+\end{figure}
+
+The discrete jump probabilities for the gamma process are obtained by discretising the gamma process.
+One property of the gamma process is that the jump probabilities are independent of the current state.
+We denote the probability of having a jump of size $j$ with rate parameter $a$ and scale parameter $b$ by $p(i; a, b)$.
+The derivation of the jump probabilities can be found in the manuscript by De Jonge on pages 51-53.
+
+\section{Markov Decision Process}
+In this section I provide the Markov Decision Process (MDP) formulation of the system as described in Section 2.
+The MDP consists of a state space, action space, transition probabilities, expected costs and a performance evaluation.
+
 \subsection{State space}
@@
 
-After selecting the maintenance action, we also have to set the production rate of each unit.
-Each unit can produce at most $m+1$ different production rates, ranging from $0$ to $1$ (maximum production).
-The possible production rates are uniformly distributed over the interval.
-If a unit is in the failed state, we cannot produce, so the production rate of this unit is set to $0$.
-For all other deteriorating levels, we can select one of the $m+1$ production rates.
  We define the set of possible production rates for unit $i$ in state $x_i$ $U(x_i)$ as follows
@@
 $$
-Hence, the 
 Let $u = (u_1, \dots, u_n)$ denote the production rate of all units in the system.
@@
 
-\subsection{Deterioration Process}
-The transition probabilities for all units are identical and independent.
-Each unit $i$ deteriorates according to a gamma process with rate $\lambda(u_i)$, which depends on the current production rate of the unit, and variance $\sigma^2$.
-The rate for a given production rate $u_i$ is given by
-$$
-\lambda(u_i) = g(u_i) \lambda_{\text{max}},
-$$
-where $\lambda_{\text{max}}$ is the rate parameter under the maximum production rate, and $g$ is some non-decreasing function.
-We assume that $g(u_i)$ as a function of the production rate $u_i$ for unit $i$ is of the following form
-$$
-g(u_i) = \beta + (1-\beta)u_i^\alpha.
-$$
-We assume that $0 \leq \beta \leq 1$ and $\alpha \geq 0$, since $g$ should be non-decreasing in $u_i$. \textbf{[[Add some information about the shape of $g$]]}
-
-For a unit $i$ the probability of having $q \geq 0$ jumps in the deterioration states for a time epoch under production rate $u_i$ is given by $p(u_i,q)$, which can be found by discretising the gamma process (see De Jonge).
+\subsection{Transition probabilities}
+For a unit $i$ the probability of having $q \geq 0$ jumps in the deterioration states for a time epoch under production rate $u_i$ and scale parameter $b$ is given by $p(q; \lambda(u_i), b)$.
 The transition probability of unit $k$ for going from state $i$ to $j$ with production rate $u_i$ and maintenance action DN for any time epoch is defined as
@@
 P_{u_k,\text{DN}}(i,j) = \begin{cases}
-p(u_k, j-i) &\text{ if } i \leq j \leq L-1, \\
+p(j-i; \lambda(u_k), b) &\text{ if } i \leq j \leq L-1, \\
 0 &\text{ if } j < i, \\
-1 - \sum_{l=0}^{L-1-i}p(u_k,l) & \text{ if } i \leq m-1  \text{ and } j = L, \\
+1 - \sum_{l=0}^{L-1-i}p(l; \lambda(u_k), b) & \text{ if } i \leq m-1  \text{ and } j = L, \\
 1 & \text{ if } i = j = L,
@@
 P_{u_k,\text{M}}(i,j) = \begin{cases}
-p(u_k, j) &\text{ if } j \leq  L-1, \\
-1 - \sum_{l=0}^{L-1}p(u_k,l) & \text{ if } j = L.
+p(j; \lambda(u_k), b)&\text{ if } j \leq  L-1, \\
+1 - \sum_{l=0}^{L-1}p(l; \lambda(u_k), b) & \text{ if } j = L.
 \end{cases}
@@
 \subsection{Performance criteria}
-As a performance criterion, we are interested in minimizing the long-run average cost per time unit.
-In this way, we can quantify the impact of adding an additional component, and find the most cost-efﬁcient maintenance policy.
+As a performance criterion, we are interested in minimising the long-run average cost per time unit.
+In this way we can quantify the maintenance costs of a system over the life cycle, and identify the benefits of being able to set the load of each unit.
 
-\section{Numerical results}
-
-\subsection{Two unit setting}
-
-\subsection{Adding additional unit}
-\begin{enumerate}
-	\item Directly compare the results for different mean rates $g$. 
-\end{enumerate}
+\section{Numerical results two-unit system}
+In this Section we will investigate the key features of the optimal maintenance and production policy for a two-unit system using the MDP as described in Section 3.
+Similar to \cite{OLDEKEIZER2018319} we select preventive maintenance costs $c_\text{pm} = 5$, set-up cost $C = 4$, and a corrective maintenance cost $c_\text{cm} = 11$.
+The rate parameter under the maximum load rate is $\lambda_{\text{max}} = 0.7$ and the scale parameter $\gamma = 1$.
+The failure level $L$ and the number of possible production rates are set to 30.
+We vary the total load $\pi$ with values 30, 40 and 50.
+For $\pi = 50$, one unit can handle all the load and for $\pi = 50$ we do not have much overcapacity in the system.
+We set $\alpha = 0.1$, such that the deterioration when a unit is not operating is low.
+We consider three different values for $\beta$:  $0$ (deterioration rate does not depend on production), $1$ (deterioration rate depends linearly on production) and $1.5$ (deterioration rate is convex in the production).
 
@@
 
-In the literature, condition-based maintenance is often implemented in the form of a threshold policy (CITE).
+In the literature, condition-based maintenance is often implemented in the form of a threshold policy (CITE), or an opportunistic threshold policy \citep{ZHANG2015176}.
+In the former policy, if the deterioration level of a unit is at or exceed a threshold level $T_r$, we perform a maintenance action.
+The opportunistic policy adds an additional step to the threshold policy: if an maintenance action is initiated, then all units that have a deterioration state higher or equal to $T_o$ are also maintained in that time period.
+The optimal threshold values for $T_r$ and $T_o$ are found using a grid search procedure.
 Due to the economic dependence (fixed cost of maintenance) and the liberty to set the production rate of the components, this policy is probably not optimal.
-For this policy, if the deterioration level of a unit is at or exceed a threshold level $T_r$, we perform a maintenance action.
-We find the optimal $T_r$ by simulating the system for a grid of values for $T_r$.
 
-Production benchmark ideas:
-\begin{enumerate}
-	\item Share production equally between all units for all states.
-	\item Random pool of $n$ units that produce equal amount.
-	\item Best 50\% produces 80\% and worst 50\% 20\%.
-\end{enumerate}
+\textbf{[[Note: I think it is unfair to compare the optimal policy to the standard threshold policy, since this policy is equal or worse than the opportunistic policy. ]]}
+
+Additionally, we need a benchmark policy for the production policy.
+Inspired by the load sharing formulation in \cite{OLDEKEIZER2018319}, we will use the following benchmark for the production policy: For all system states determine the optimal selection of functioning units, where all functioning units share the load uniformly.
+The optimal selection is found by minimising the long-run average  cost per time unit.
+
+Table (???) presents the
 
@@
 
+\section{Numerical results three-unit system}
+
+\section{Conclusion}
+% Summary
+
+% Future research
+In this paper, I do not include dependence between the capacity of units and its deterioration state.
+
+\bibliographystyle{IEEEtranSN} 
+\bibliography{mylib}
+
 \appendix
@@
 \subsection{Discretising the gamma process}
+See Maintenance planning and optimisation literature of Bram de Jonge.
+
+\end{document}

### 2018-05-30 general_figures.R

@@ -1,0 +1,27 @@
+library(tidyverse)
+library(reshape2)
+
+g = function(u, a, b){ b + (1 - b) * u^a}
+g = Vectorize(g, "u")
+
+u_range = seq(0, 1, length.out=100)
+
+alphas = c(0, 0.5, 1, 1.5)
+betas = c(0.1)
+
+data = NULL
+for (alpha in alphas) {
+  for (beta in betas) {
+    data = data %>%
+      bind_rows(tibble(u = u_range, g = g(u_range, alpha, beta), 
+                       setting = paste("a = ", alpha, ", b = ", beta, sep="")))
+  }
+}
+
+ggplot(data, aes(x = u, y = g, colour = setting)) + geom_line() + 
+  theme_bw() + theme(panel.grid = element_blank()) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0.1)) +
+  labs(x="production", y="g")
+
+ggsave("report/figures/g.pdf", width=7, height=5)

### 2018-05-30 src/mdp/applied_policy.cpp

@@ -8,8 +8,8 @@
  * Or do we determine the thresholds of our maintenancePolicy?
  */
 
-num_t MDP::applied_policy(std::vector<std::array<int, 2>> maintenancePolicy,
-                          std::vector<std::array<int, 2>> productionPolicy,
+num_t MDP::applied_policy(std::vector<std::array<int, 2> > maintenancePolicy,
+                          std::vector<std::array<int, 2> > productionPolicy,
                           int simTime) const
 {
     double totalCosts = 0;

### 2018-05-30 src/mdp/mdp.h

@@ -20,8 +20,8 @@ class MDP
     std::vector<num_t> d_expectedCost;
 
     // Save optimal actions
-    std::vector<std::array<int, 2>> d_optimalMaintenance;
-    std::vector<std::array<int, 2>> d_optimalProduction;
+    std::vector<std::array<int, 2> > d_optimalMaintenance;
+    std::vector<std::array<int, 2> > d_optimalProduction;
     // std::vector<std::array<int, 3>> d_optimalMaintenance;
     // std::vector<std::array<int, 3>> d_optimalProduction;
 
@@ -57,8 +57,8 @@ class MDP
         void    solve();
         void    write_solution() const;
         void    print_all_info() const;
-        num_t   applied_policy(std::vector<std::array<int, 2>> maintenancePolicy,
-                               std::vector<std::array<int, 2>> productionPolicy,
+        num_t   applied_policy(std::vector<std::array<int, 2> > maintenancePolicy,
+                               std::vector<std::array<int, 2> > productionPolicy,
                                int simTime) const;
         
         size_t get_id() const;

### 2018-05-30 create_experiments.R
Implemented heuristic

@@ -6,9 +6,9 @@ library(tidyverse)
 
 grid_size = 10
 
-base_experiment = tibble(ccm = 1.0,
-                         cpm = 0.2,
-                         C = 0.5,
+base_experiment = tibble(ccm = 11.0,
+                         cpm = 5.0,
+                         C = 4.0,
                          nStates = 30,
                          dL = 1.0,
                          pi = 40,
@@ -33,6 +33,16 @@ create_experiments = function(variableName, values) {
   
   return(new_experiments)
 }
+
+# Set pi
+pi_options = seq(30, 40, 50)
+experiments = experiments %>%
+  bind_rows(create_experiments("pi", pi_options))
+
+# Set alpha
+alpha_options = seq(0, 1, 1.5)
+experiments = experiments %>%
+  bind_rows(create_experiments("alpha", pi_options))
 
 # Variance experiments
 var_options = seq(0.5, 5, length.out=grid_size)

### 2018-05-30 src/mdp/applied_policy.cpp

@@ -12,19 +12,48 @@ num_t MDP::applied_policy(std::vector<std::array<int, 2> > maintenancePolicy,
                           std::vector<std::array<int, 2> > productionPolicy,
                           int simTime) const
 {
-    double totalCosts = 0;
-    double detState1 = 0;
-    double detState2 = 0;
+    // Initialise random generator
+    random_device rd;
+    mt19937 gen(rd());
+
+    num_t totalCosts = 0;
+    num_t detState1 = 0;
+    num_t detState2 = 0;
+    vector<vector<discrete_distribution<>>> jumpProbabilities = distribution_sampler();
 
     // Run simulation for simTime time units
     for (int t = 0; t != simTime; ++t)
     {
-        // Check if we perform maintenance
+        // Check if we perform maintenance on unit 1
+        if (maintenancePolicy[hash(detState1, detState2)][1])
+        {
+            detState1 = 0;
+            if (detState1 == (d_nStates-1))
+                totalCosts += d_ccm;
+            else
+                totalCosts += d_cpm;
+        }
 
-        // Set production rate
+        // Check if we perform maintenance on unit 2
+        if (maintenancePolicy[hash(detState1, detState2)][2])
+        {
+            detState2 = 0;
+            if (detState2 == (d_nStates-1))
+                totalCosts += d_ccm;
+            else
+                totalCosts += d_cpm;
+        }
 
-        // Determine deterioration (DISCRETE OR CONTINUOUS)
+        // Check if we have to pay fixed costs, i.e. at least one
+        // maintenance action.
+        if (maintenancePolicy[hash(detState1, detState2)][1] or maintenancePolicy[hash(detState1, detState2)][2])
+            totalCosts += d_C;
 
+        // Determine deterioration
+        int prodRate1 = productionPolicy[hash(detState1, detState2)][0];
+        int prodRate2 = productionPolicy[hash(detState1, detState2)][0];
+        detState1 += jumpProbabilities[prodRate1][detState1](gen);
+        detState2 += jumpProbabilities[prodRate2][detState2](gen);
     }
 
     // Mean cost per time unit is total cost divided by length of time

### 2018-05-30 src/mdp/converged_heuristic.cpp

@@ -1,0 +1,8 @@
+#include "mdp.ih"
+
+bool MDP::converged_heuristic() const
+{
+    num_t const span = bs::span(d_valueBHeuristic, d_valueAHeuristic);
+    
+    return (span < d_eps);
+}

### 2018-05-30 src/mdp/cost_heuristic.cpp

@@ -1,0 +1,36 @@
+#include "mdp.ih"
+
+/*
+ * Obtain optimal values for the heuristic and return the average cost.
+ * Heuristic consists of two parts:
+ *  1) Maintenaince heuristic
+ *  2) Production heuristic
+ */
+
+num_t MDP::cost_heuristic(bool mHeuristic, bool pHeuristic)
+{
+    num_t minCosts = numeric_limits<num_t>::max();
+
+    if (mHeuristic)
+    {
+        for (int T_r = d_nStates / 2; T_r != d_nStates; ++T_r)
+        {
+            for (int T_o = 0; T_o != T_r; ++T_o)
+            {
+                // Evaluate policy
+                solve_heuristic(mHeuristic, pHeuristic, T_r, T_o);
+                num_t const tmpCosts = bs::approximate_gain(d_valueBHeuristic, d_valueAHeuristic);
+
+                if (tmpCosts < minCosts)
+                    minCosts = tmpCosts;
+            }
+        }
+    } else 
+    {
+        // Evaluate policy
+        solve_heuristic(mHeuristic, pHeuristic, -1, -1);
+        minCosts = bs::approximate_gain(d_valueBHeuristic, d_valueAHeuristic);   
+    }
+
+    return minCosts;
+}

### 2018-05-30 src/mdp/distribution_sampler.cpp

@@ -1,0 +1,34 @@
+#include "mdp.ih"
+
+/*
+ * Create a discrete distribution for every production rate in
+ * every deterioration state.
+ */
+
+vector<vector<discrete_distribution<>>> MDP::distribution_sampler() const
+{
+    vector<vector<discrete_distribution<>>> jumpProbabilities;
+    jumpProbabilities.resize(d_nRates);
+
+    for (int sdx = 0; sdx != d_nStates; ++sdx)
+    {
+        for (int pdx = 0; pdx != d_nRates; ++pdx)
+        {
+            jumpProbabilities[pdx].resize(d_nStates);
+
+            // Maximum number of jumps to non-failed states
+            int maxJumps = (d_nStates-1) - sdx;
+            vector<double> jumpProbability(maxJumps+1, 0);
+            
+            // Add all jumps to non-failed states
+            for (int jump = 0; jump != maxJumps; ++jump)
+                jumpProbability[jump] = d_prop1[pdx][jump];
+            // Jump to failed state
+            jumpProbability[maxJumps] = d_prop2[pdx][sdx];
+
+            jumpProbabilities[pdx][sdx] = discrete_distribution<>(begin(jumpProbability), end(jumpProbability));
+        }
+    }
+
+    return jumpProbabilities;
+}

### 2018-05-30 src/mdp/exp_cost_heuristic.cpp

@@ -1,0 +1,39 @@
+#include "mdp.ih"
+
+/*
+    For every state we calculate the new value 
+    to obtain V_t+1
+*/
+
+void MDP::exp_cost_heuristic1()
+{
+    int idxState = 0;
+
+    for (int state1 = 0; state1 != d_nStates; ++state1)
+    {
+        for (int state2 = 0; state2 != d_nStates; ++state2) 
+        {
+            d_expectedCostHeuristic[idxState] = expected_cost_heuristic1(state1, state2, idxState);
+            ++idxState;
+        }
+    }    
+}
+
+/*
+    For every state we calculate the new value 
+    to obtain V_t+1 (without production heuristic)
+*/
+
+void MDP::exp_cost_heuristic2()
+{
+    int idxState = 0;
+
+    for (int state1 = 0; state1 != d_nStates; ++state1)
+    {
+        for (int state2 = 0; state2 != d_nStates; ++state2) 
+        {
+            d_expectedCostHeuristic[idxState] = expected_cost_heuristic2(state1, state2, idxState);
+            ++idxState;
+        }
+    }    
+}

### 2018-05-30 src/mdp/expected_costs_heuristic.cpp

@@ -1,0 +1,79 @@
+#include "mdp.ih"
+
+/*
+ * Expected cost using production heuristic
+ */
+
+num_t MDP::expected_cost_heuristic1(int state1, int state2, int idxState)
+{
+    num_t bestVal = numeric_limits<num_t>::max();
+    int bestProd1 = -1;
+    int bestProd2 = -1;
+
+    // Check if we should do one working unit
+    if (d_pi <= d_nRates)
+    {
+        bestVal = expected_costs(state1, state2, d_pi, 0);
+        bestProd1 = d_pi;
+        bestProd2 = 0;
+
+        num_t const tmpVal2 = expected_costs(state1, state2, 0, d_pi);
+
+        if (tmpVal2 < bestVal)
+        {
+            bestVal = tmpVal2;
+            bestProd1 = 0;
+            bestProd2 = d_pi;
+        }
+    }
+
+    // Share load
+    int pdx1 = d_pi / 2;
+    int pdx2 = d_pi - pdx1;
+    num_t tmpVal3 = expected_costs(state1, state2, pdx1, pdx2);
+
+    if (tmpVal3 < bestVal)
+    {
+        bestVal = tmpVal3;
+        bestProd1 = pdx1;
+        bestProd2 = pdx2;
+    }
+
+    d_optimalProductionHeuristic[idxState][0] = bestProd1;
+    d_optimalProductionHeuristic[idxState][1] = bestProd2;
+    
+    return bestVal;
+}
+
+/*
+ * Expected cost using production heuristic
+ */
+
+num_t MDP::expected_cost_heuristic2(int state1, int state2, int idxState)
+{
+    num_t bestVal = numeric_limits<num_t>::max();
+    int bestProd1 = -1;
+    int bestProd2 = -1;
+
+    // Select minimum costs for all possible production states
+    for (int pdx1 = (d_pi - d_nRates + 1); pdx1 != d_nRates; ++pdx1)
+    {
+        int pdx2 = d_pi - pdx1;
+        if (not feasible_production(state1, state2, pdx1, pdx2))
+            continue;
+
+        num_t const tmpVal = expected_costs(state1, state2, pdx1, pdx2);
+
+        if (tmpVal < bestVal - gg_eps) 
+        {
+            bestVal = tmpVal;
+            bestProd1 = pdx1;
+            bestProd2 = pdx2;
+        }
+    }
+
[…]

### 2018-05-30 src/mdp/mdp.cpp

@@ -22,7 +22,12 @@ MDP::MDP(Init const &init, bool jumpHeuristic)
     d_jumpHeuristic(jumpHeuristic),
     d_maxJumps(init.d_maxJumps),
     d_terminalCosts(init.d_terminalCosts),
-    d_eps(init.d_eps)
+    d_eps(init.d_eps),
+    d_optimalMaintenanceHeuristic(vector<array<int, 2>>(init.d_nStates * init.d_nStates, array<int, 2>())),
+    d_optimalProductionHeuristic(vector<array<int, 2>>(init.d_nStates * init.d_nStates, array<int, 2>())),
+    d_valueAHeuristic(vector<num_t>(init.d_nStates * init.d_nStates, 0)),
+    d_valueBHeuristic(vector<num_t>(init.d_nStates * init.d_nStates, 0)),
+    d_expectedCostHeuristic(vector<num_t>(init.d_nStates * init.d_nStates, 0))
 {
     if (d_pi < d_nRates || d_pi > (2 * d_nRates))
         throw string("Pi should be between nRates and 2*nRates");

### 2018-05-30 src/mdp/mdp.h

@@ -4,6 +4,7 @@
 #include <vector>
 #include <array>
 #include <tuple>
+#include <random>
 
 #include "../init/init.h"
 
@@ -50,16 +51,24 @@ class MDP
     // Value iteration parameter
     num_t const d_eps;
 
+    // Heuristic
+    std::vector<std::array<int, 2> > d_optimalMaintenanceHeuristic;
+    std::vector<std::array<int, 2> > d_optimalProductionHeuristic;
+
+    std::vector<num_t> d_valueAHeuristic;
+    std::vector<num_t> d_valueBHeuristic;
+
+    std::vector<num_t> d_expectedCostHeuristic;
+
+
     public:
         MDP() = delete;
         MDP(Init const &init, bool jumpHeuristic = false);
 
         void    solve();
-        void    write_solution() const;
+        void    write_solution();
         void    print_all_info() const;
-        num_t   applied_policy(std::vector<std::array<int, 2> > maintenancePolicy,
-                               std::vector<std::array<int, 2> > productionPolicy,
-                               int simTime) const;
+        void    print_policy(std::vector<std::array<int, 2>> const &policy) const;
         
         size_t get_id() const;
 
@@ -69,6 +78,27 @@ class MDP
         void    exp_cost();
         bool    converged() const;
         num_t   direct_costs(int state, int mainAction) const;
+
+        // Heuristic functions
+        num_t   cost_heuristic(bool mHeuristic, bool pHeuristic);
+        num_t   applied_policy(std::vector<std::array<int, 2> > maintenancePolicy,
+                               std::vector<std::array<int, 2> > productionPolicy,
+                               int simTime) const;
+        void    set_maintenance_matrix(int T_r, int T_o);
+        num_t   expected_cost_heuristic1(int state1, int state2, int idxState);
+        num_t   expected_cost_heuristic2(int state1, int state2, int idxState);
+        void    exp_cost_heuristic1();
+        void    exp_cost_heuristic2();
+        num_t   value_costs_heuristic1(int state1, int state2, int idxState);
+        num_t   value_costs_heuristic2(int state1, int state2, int idxState);
+        num_t   value_costs_heuristic3(int state1, int state2, int idxState);
+        void    value_cost_heuristic1();
+        void    value_cost_heuristic2();
+        void    value_cost_heuristic3();
+        void    solve_heuristic(bool mHeuristic, bool pHeuristic, int T_r, int T_o);
+        bool    converged_heuristic() const;
+
[…]

### 2018-05-30 src/mdp/print_policy.cpp

@@ -1,0 +1,23 @@
+#include "mdp.ih"
+
+/*
+ *  Print policy in wide format
+ */
+
+void MDP::print_policy(vector<array<int, 2>> const &policy) const
+{
+    // Header
+    cout << setw(6) << "State ";
+    for (int sdx2 = 0; sdx2 != d_nStates; ++sdx2)
+        cout << setw(6) << sdx2 << ' ';
+    cout << endl;
+
+    // Values
+    for (int sdx1 = 0; sdx1 != d_nStates; ++sdx1)
+    {
+        cout << setw(6) << sdx1 << ' ';
+        for (int sdx2 = 0; sdx2 != d_nStates; ++sdx2)
+            cout << setw(6) << string("(") + to_string(policy[hash(sdx1, sdx2)][0]) + string(",") + to_string(policy[hash(sdx1, sdx2)][1]) + string(")") << ' ';
+        cout << endl;
+    }
+}

### 2018-05-30 src/mdp/set_maintenance_matrix.cpp

@@ -1,0 +1,41 @@
+#include "mdp.ih"
+
+/*
+ * Return matrix with opportunistic maintenance heuristic.
+ */
+
+void MDP::set_maintenance_matrix(int T_r, int T_o)
+{
+    for (int sdx1 = 0; sdx1 != d_nStates; ++sdx1)
+    {
+        for (int sdx2 = 0; sdx2 != d_nStates; ++sdx2)
+        {
+            if (sdx1 >= T_r)
+            {
+                d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][0] = 1;
+                if (sdx2 >= T_o)
+                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][1] = 1;
+                else
+                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][1] = 0;
+            } else if (sdx1 >= T_o)
+            {
+                if (sdx2 >= T_r)
+                {
+                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][0] = 1;
+                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][1] = 1;
+                } else 
+                {
+                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][0] = 0;
+                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][1] = 0;
+                }
+            } else 
+            {
+                d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][0] = 0;
+                if (sdx2 >= T_r)
+                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][1] = 1;
+                else
+                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][1] = 0;
+            }
+        }
+    }
+}

### 2018-05-30 src/mdp/solve_heuristic.cpp

@@ -1,0 +1,40 @@
+#include "mdp.ih"
+
+void MDP::solve_heuristic(bool mHeuristic, bool pHeuristic, int T_r, int T_o)
+{
+    // If maintenance heuristic determine maintenance policy
+    if (mHeuristic)
+        set_maintenance_matrix(T_r, T_o);
+
+    d_valueBHeuristic = vector<num_t>(d_nStates * d_nStates, 0);
+
+    // Use all jump probabilities
+    do
+    {
+        // Set new values to old values
+        d_valueAHeuristic = d_valueBHeuristic;
+
+        // Compute the expected cost for every state and optimal production
+        if (pHeuristic)
+            exp_cost_heuristic1();
+        else
+            exp_cost_heuristic2();
+
+        // Obtain new value
+        if (pHeuristic and mHeuristic)
+        {  
+            value_cost_heuristic3();
+        } else if (mHeuristic)
+        {
+            value_cost_heuristic1();
+        }
+        else if (pHeuristic) 
+        {
+            value_cost_heuristic2();
+        } else
+        {
+            throw(string("Specify heuristic"));
+        }
+
+    } while(not converged_heuristic());
+}

### 2018-05-30 src/mdp/value_cost_heuristic.cpp

@@ -1,0 +1,43 @@
+#include "mdp.ih"
+
+void MDP::value_cost_heuristic1()
+{
+    int idxState = 0;
+    
+    for (int state1 = 0; state1 != d_nStates; ++state1)
+    {
+        for (int state2 = 0; state2 != d_nStates; ++state2)
+        {
+            d_valueBHeuristic[idxState] = value_costs_heuristic1(state1, state2, idxState);
+            ++idxState;
+        }
+    }
+}
+
+void MDP::value_cost_heuristic2()
+{
+    int idxState = 0;
+    
+    for (int state1 = 0; state1 != d_nStates; ++state1)
+    {
+        for (int state2 = 0; state2 != d_nStates; ++state2)
+        {
+            d_valueBHeuristic[idxState] = value_costs_heuristic2(state1, state2, idxState);
+            ++idxState;
+        }
+    }
+}
+
+void MDP::value_cost_heuristic3()
+{
+    int idxState = 0;
+    
+    for (int state1 = 0; state1 != d_nStates; ++state1)
+    {
+        for (int state2 = 0; state2 != d_nStates; ++state2)
+        {
+            d_valueBHeuristic[idxState] = value_costs_heuristic3(state1, state2, idxState);
+            ++idxState;
+        }
+    }
+}

### 2018-05-30 src/mdp/value_costs_heuristic.cpp

@@ -1,0 +1,63 @@
+#include "mdp.ih"
+
+/*
+ * Value costs for using heuristic for only maintenance 
+ */
+num_t MDP::value_costs_heuristic1(int state1, int state2, int idxState)
+{
+    int adx1 = d_optimalMaintenanceHeuristic[idxState][0];
+    int adx2 = d_optimalMaintenanceHeuristic[idxState][1];
+    int s1 = (adx1 == 0) ? state1 : 0;
+    int s2 = (adx2 == 0) ? state2 : 0;
+
+    return direct_cost(state1, state2, adx1, adx2) + d_expectedCost[hash(s1, s2)];
+}
+
+/*
+ * Value costs for using heuristic for only production
+ */
+num_t MDP::value_costs_heuristic2(int state1, int state2, int idxState)
+{
+    num_t bestVal = numeric_limits<num_t>::max();
+    int bestMain1 = -1;
+    int bestMain2 = -1;
+
+    for (int adx1 = 0; adx1 != 2; ++adx1)
+    {
+        int s1 = (adx1 == 0) ? state1 : 0;
+        for (int adx2 = 0; adx2 != 2; ++adx2)
+        {
+            if (not feasible_maintenance(state1, state2, adx1, adx2))
+                continue;
+
+            int s2 = (adx2 == 0) ? state2 : 0;
+
+            num_t const tmpVal = direct_cost(state1, state2, adx1, adx2) + d_expectedCostHeuristic[hash(s1, s2)];
+            if (tmpVal < bestVal - gg_eps)
+            {
+                bestVal = tmpVal;
+                bestMain1 = adx1;
+                bestMain2 = adx2;
+            }
+        }
+    }
+
+    d_optimalMaintenanceHeuristic[idxState][0] = bestMain1;
+    d_optimalMaintenanceHeuristic[idxState][1] = bestMain2;
+    
+    return bestVal;  
+}
+
+/*
+ * Value costs for using heuristic for both maintenance
+ * and production
+ */
+num_t MDP::value_costs_heuristic3(int state1, int state2, int idxState)
+{
+    int adx1 = d_optimalMaintenanceHeuristic[idxState][0];
+    int adx2 = d_optimalMaintenanceHeuristic[idxState][1];
+    int s1 = (adx1 == 0) ? state1 : 0;
+    int s2 = (adx2 == 0) ? state2 : 0;
+
+    return direct_cost(state1, state2, adx1, adx2) + d_expectedCostHeuristic[hash(s1, s2)];
+}

### 2018-05-30 src/mdp/write_solution.cpp

@@ -8,7 +8,7 @@
  * production in every state for every unit.
  */
 
-void MDP::write_solution() const
+void MDP::write_solution()
 {
     // Save the optimal policy
     ofstream outputFile;
@@ -42,4 +42,10 @@ void MDP::write_solution() const
     outputFile.open("output/approximate_gain" + to_string(d_id));
     outputFile << bs::approximate_gain(d_valueB, d_valueA) << endl;
     outputFile.close();
+
+    // Obtain heuristic value
+    num_t heuristicCost = cost_heuristic(false, true);
+    print_policy(d_optimalMaintenanceHeuristic);
+    print_policy(d_optimalProductionHeuristic);
+    cout << heuristicCost << endl;
 }

### 2018-05-30 src/run.cpp

@@ -20,10 +20,10 @@ try
         Init init{1,        // ccm
                   0.2,      // cpm
                   0.5,      // C
-                  50,       // nStates
+                  20,       // nStates
                   1,        // dL
-                  60,       // pi
-                  50,       // nRates
+                  30,       // pi
+                  20,       // nRates
                   1,        // mu
                   2,        // var
                   1,        // dT

### 2018-05-30 src/mdp/expected_cost_heuristic.cpp
Updated heuristic

@@ -1,0 +1,79 @@
+#include "mdp.ih"
+
+/*
+ * Expected cost using production heuristic
+ */
+
+num_t MDP::expected_cost_heuristic1(int state1, int state2, int idxState)
+{
+    num_t bestVal = numeric_limits<num_t>::max();
+    int bestProd1 = -1;
+    int bestProd2 = -1;
+
+    // Check if we should do one working unit
+    if (d_pi <= d_nRates)
+    {
+        bestVal = expected_costs_heuristic(state1, state2, d_pi, 0);
+        bestProd1 = d_pi;
+        bestProd2 = 0;
+
+        num_t const tmpVal2 = expected_costs_heuristic(state1, state2, 0, d_pi);
+
+        if (tmpVal2 < bestVal)
+        {
+            bestVal = tmpVal2;
+            bestProd1 = 0;
+            bestProd2 = d_pi;
+        }
+    }
+
+    // Share load
+    int pdx1 = d_pi / 2;
+    int pdx2 = d_pi - pdx1;
+    num_t tmpVal3 = expected_costs_heuristic(state1, state2, pdx1, pdx2);
+
+    if (tmpVal3 < bestVal)
+    {
+        bestVal = tmpVal3;
+        bestProd1 = pdx1;
+        bestProd2 = pdx2;
+    }
+
+    d_optimalProductionHeuristic[idxState][0] = bestProd1;
+    d_optimalProductionHeuristic[idxState][1] = bestProd2;
+    
+    return bestVal;
+}
+
+/*
+ * Expected cost using production heuristic
+ */
+
+num_t MDP::expected_cost_heuristic2(int state1, int state2, int idxState)
+{
+    num_t bestVal = numeric_limits<num_t>::max();
+    int bestProd1 = -1;
+    int bestProd2 = -1;
+
+    // Select minimum costs for all possible production states
+    for (int pdx1 = (d_pi - d_nRates + 1); pdx1 != d_nRates; ++pdx1)
+    {
+        int pdx2 = d_pi - pdx1;
+        if (not feasible_production(state1, state2, pdx1, pdx2))
+            continue;
+
+        num_t const tmpVal = expected_costs_heuristic(state1, state2, pdx1, pdx2);
+
+        if (tmpVal < bestVal - gg_eps) 
+        {
+            bestVal = tmpVal;
+            bestProd1 = pdx1;
+            bestProd2 = pdx2;
+        }
+    }
+
[…]

### 2018-05-30 src/mdp/expected_costs_heuristic.cpp

@@ -1,79 +1,46 @@
 #include "mdp.ih"
 
 /*
- * Expected cost using production heuristic
+ *  Calculate the expected cost for a given state and action.
+ *  Loop over all possible jumps to states
+ * HEURISTIC
  */
 
-num_t MDP::expected_cost_heuristic1(int state1, int state2, int idxState)
+num_t MDP::expected_costs_heuristic(int state1, int state2, int prodRate1, int prodRate2) const
 {
-    num_t bestVal = numeric_limits<num_t>::max();
-    int bestProd1 = -1;
-    int bestProd2 = -1;
+    num_t costs = 0;
 
-    // Check if we should do one working unit
-    if (d_pi <= d_nRates)
+    // Check if we do all jumps or selection.
+    int maxJump1 = d_nStates-state1-1;
+    int maxJump2 = d_nStates-state2-1;
+
+    if (d_jumpHeuristic) 
     {
-        bestVal = expected_costs(state1, state2, d_pi, 0);
-        bestProd1 = d_pi;
-        bestProd2 = 0;
+        maxJump1 = min(maxJump1, d_maxJumps[prodRate1]);
+        maxJump2 = min(maxJump2, d_maxJumps[prodRate2]);
+    } 
 
-        num_t const tmpVal2 = expected_costs(state1, state2, 0, d_pi);
+    // Loop over all jump probabilities from 'state1' to 'd_nStates' and obtain costs.
+    int idxState = hash(state1, 0);
+    for (int sdx1 = 0; sdx1 != maxJump1; ++sdx1) 
+    {
+        num_t prop1 = d_prop1[prodRate1][sdx1];
+        idxState += state2;
 
-        if (tmpVal2 < bestVal)
-        {
-            bestVal = tmpVal2;
-            bestProd1 = 0;
-            bestProd2 = d_pi;
-        }
+        for (int sdx2 = 0; sdx2 != maxJump2; ++sdx2)
+            costs += prop1 * d_prop1[prodRate2][sdx2] * d_valueAHeuristic[idxState++];
+        costs += prop1 * d_prop2[prodRate2][state2] * d_valueAHeuristic[idxState++];
     }
 
-    // Share load
-    int pdx1 = d_pi / 2;
-    int pdx2 = d_pi - pdx1;
-    num_t tmpVal3 = expected_costs(state1, state2, pdx1, pdx2);
+    // Add costs when unit 1 goes to failed state;
[…]

### 2018-05-30 src/mdp/mdp.h

@@ -85,6 +85,7 @@ class MDP
                                std::vector<std::array<int, 2> > productionPolicy,
                                int simTime) const;
         void    set_maintenance_matrix(int T_r, int T_o);
+        num_t   expected_costs_heuristic(int state1, int state2, int prodRate1, int prodRate2) const;
         num_t   expected_cost_heuristic1(int state1, int state2, int idxState);
         num_t   expected_cost_heuristic2(int state1, int state2, int idxState);
         void    exp_cost_heuristic1();

### 2018-05-30 src/mdp/write_solution.cpp

@@ -44,7 +44,7 @@ void MDP::write_solution()
     outputFile.close();
 
     // Obtain heuristic value
-    num_t heuristicCost = cost_heuristic(false, true);
+    num_t heuristicCost = cost_heuristic(true, true);
     print_policy(d_optimalMaintenanceHeuristic);
     print_policy(d_optimalProductionHeuristic);
     cout << heuristicCost << endl;

### 2018-05-31 create_experiments.R
Heuristic finished

@@ -9,10 +9,10 @@ grid_size = 10
 base_experiment = tibble(ccm = 11.0,
                          cpm = 5.0,
                          C = 4.0,
-                         nStates = 30,
+                         nStates = 31,
                          dL = 1.0,
                          pi = 40,
-                         nRates = 30,
+                         nRates = 31,
                          mu = 1.0,
                          var = 2.0,
                          dT = 1.0,
@@ -35,24 +35,24 @@ create_experiments = function(variableName, values) {
 }
 
 # Set pi
-pi_options = seq(30, 40, 50)
+pi_options = c(20, 30, 40, 58)
 experiments = experiments %>%
   bind_rows(create_experiments("pi", pi_options))
 
 # Set alpha
-alpha_options = seq(0, 1, 1.5)
+alpha_options = c(0, 1, 1.5)
 experiments = experiments %>%
-  bind_rows(create_experiments("alpha", pi_options))
+  bind_rows(create_experiments("alpha", alpha_options))
 
-# Variance experiments
-var_options = seq(0.5, 5, length.out=grid_size)
-experiments = experiments %>%
-  bind_rows(create_experiments("var", var_options))
-
-# Fixed cost effect
-fixed_cost_options = seq(0, 3, length.out=grid_size)
-experiments = experiments %>%
-  bind_rows(create_experiments("C", fixed_cost_options))
+# # Variance experiments
+# var_options = seq(0.5, 5, length.out=grid_size)
+# experiments = experiments %>%
+#   bind_rows(create_experiments("var", var_options))
+# 
+# # Fixed cost effect
+# fixed_cost_options = seq(0, 3, length.out=grid_size)
+# experiments = experiments %>%
+#   bind_rows(create_experiments("C", fixed_cost_options))
 
 
 #### Output the experiments table ####

### 2018-05-31 src/mdp/cost_heuristic.cpp

@@ -15,7 +15,7 @@ num_t MDP::cost_heuristic(bool mHeuristic, bool pHeuristic)
     {
         for (int T_r = d_nStates / 2; T_r != d_nStates; ++T_r)
         {
-            for (int T_o = 0; T_o != T_r; ++T_o)
+            for (int T_o = 0; T_o <= T_r; ++T_o)
             {
                 // Evaluate policy
                 solve_heuristic(mHeuristic, pHeuristic, T_r, T_o);

### 2018-05-31 src/mdp/expected_cost.cpp

@@ -16,7 +16,10 @@ num_t MDP::expected_cost(int state1, int state2, int idxState)
     // Select minimum costs for all possible production states
     for (int pdx1 = (d_pi - d_nRates + 1); pdx1 != d_nRates; ++pdx1)
     {
-        int pdx2 = d_pi - pdx1;
+        int pdx2 = 0;
+        if (pdx1 < d_pi)
+            pdx2 = d_pi - pdx1;
+
         if (not feasible_production(state1, state2, pdx1, pdx2))
             continue;
 

### 2018-05-31 src/mdp/expected_cost_heuristic.cpp

@@ -58,7 +58,10 @@ num_t MDP::expected_cost_heuristic2(int state1, int state2, int idxState)
     // Select minimum costs for all possible production states
     for (int pdx1 = (d_pi - d_nRates + 1); pdx1 != d_nRates; ++pdx1)
     {
-        int pdx2 = d_pi - pdx1;
+        int pdx2 = 0;
+        if (pdx1 < d_pi)
+            pdx2 = d_pi - pdx1;
+        
         if (not feasible_production(state1, state2, pdx1, pdx2))
             continue;
 

### 2018-05-31 src/mdp/feasible_production.cpp

@@ -6,6 +6,9 @@
 
 bool MDP::feasible_production(int state1, int state2, int prodRate1, int prodRate2) const
 {
+    if (prodRate1 + prodRate2 != d_pi)
+        return false;
+    
     return not (((state1 == (d_nStates - 1)) and (prodRate1 > 0)) or ((state2 == (d_nStates - 1)) and (prodRate2 > 0)));
 }
 

### 2018-05-31 src/mdp/mdp.cpp

@@ -29,8 +29,8 @@ MDP::MDP(Init const &init, bool jumpHeuristic)
     d_valueBHeuristic(vector<num_t>(init.d_nStates * init.d_nStates, 0)),
     d_expectedCostHeuristic(vector<num_t>(init.d_nStates * init.d_nStates, 0))
 {
-    if (d_pi < d_nRates || d_pi > (2 * d_nRates))
-        throw string("Pi should be between nRates and 2*nRates");
+    if (d_pi > 2 * (d_nRates-1))
+        throw string("Error: pi should be feasible for units");
 
     ++s_id;
 }

### 2018-05-31 src/mdp/mdp.h

@@ -65,10 +65,12 @@ class MDP
         MDP() = delete;
         MDP(Init const &init, bool jumpHeuristic = false);
 
-        void    solve();
-        void    write_solution();
-        void    print_all_info() const;
-        void    print_policy(std::vector<std::array<int, 2>> const &policy) const;
+        void        solve();
+        void        write_solution();
+        void        print_all_info() const;
+        void        print_policy(std::vector<std::array<int, 2>> const &policy) const;
+        std::string print_policy_file(std::vector<std::array<int, 2>> const &maintenancePolicy,
+                                      std::vector<std::array<int, 2>> const &productionPolicy) const;
         
         size_t get_id() const;
 

### 2018-05-31 src/mdp/print_policy.cpp

@@ -20,4 +20,5 @@ void MDP::print_policy(vector<array<int, 2>> const &policy) const
             cout << setw(6) << string("(") + to_string(policy[hash(sdx1, sdx2)][0]) + string(",") + to_string(policy[hash(sdx1, sdx2)][1]) + string(")") << ' ';
         cout << endl;
     }
+    cout << endl << endl;
 }

### 2018-05-31 src/mdp/print_policy_file.cpp

@@ -1,0 +1,25 @@
+#include "mdp.ih"
+
+string MDP::print_policy_file(vector<array<int, 2>> const &maintenancePolicy,
+                              vector<array<int, 2>> const &productionPolicy) const 
+{
+    ostringstream os;
+
+    // Define header of the file (space separated)
+    os << "state1 state2 maintenance1 maintenance2 production1 production2" << endl;
+
+    // Loop over all states
+    int idxState = 0;
+    for (int sdx1 = 0; sdx1 != d_nStates; ++sdx1)
+    {
+        for (int sdx2 = 0; sdx2 != d_nStates; ++sdx2)
+        {
+            os << sdx1 << ' ' << sdx2 << ' ';
+            os << maintenancePolicy[idxState][0] << ' ' << maintenancePolicy[idxState][1] << ' ';
+            os << productionPolicy[idxState][0] << ' ' << productionPolicy[idxState][1] << endl;
+            ++idxState;
+        }
+    }
+
+    return os.str();
+}

### 2018-05-31 src/mdp/value_costs_heuristic.cpp

@@ -10,7 +10,7 @@ num_t MDP::value_costs_heuristic1(int state1, int state2, int idxState)
     int s1 = (adx1 == 0) ? state1 : 0;
     int s2 = (adx2 == 0) ? state2 : 0;
 
-    return direct_cost(state1, state2, adx1, adx2) + d_expectedCost[hash(s1, s2)];
+    return direct_cost(state1, state2, adx1, adx2) + d_expectedCostHeuristic[hash(s1, s2)];
 }
 
 /*

### 2018-05-31 src/mdp/write_solution.cpp

@@ -13,22 +13,7 @@ void MDP::write_solution()
     // Save the optimal policy
     ofstream outputFile;
     outputFile.open("output/optimal_policy" + to_string(d_id) + ".csv");
-
-    // Define header of the file (space separated)
-    outputFile << "state1 state2 maintenance1 maintenance2 production1 production2" << endl;
-
-    // Loop over all states
-    int idxState = 0;
-    for (int sdx1 = 0; sdx1 != d_nStates; ++sdx1)
-    {
-        for (int sdx2 = 0; sdx2 != d_nStates; ++sdx2)
-        {
-            outputFile << sdx1 << ' ' << sdx2 << ' ';
-            outputFile << d_optimalMaintenance[idxState][0] << ' ' << d_optimalMaintenance[idxState][1] << ' ';
-            outputFile << d_optimalProduction[idxState][0] << ' ' << d_optimalProduction[idxState][1] << endl;
-            ++idxState;
-        }
-    }
+    outputFile << print_policy_file(d_optimalMaintenance, d_optimalProduction);
     outputFile.close();
 
     // Save optimal values
@@ -43,9 +28,28 @@ void MDP::write_solution()
     outputFile << bs::approximate_gain(d_valueB, d_valueA) << endl;
     outputFile.close();
 
-    // Obtain heuristic value
-    num_t heuristicCost = cost_heuristic(true, true);
-    print_policy(d_optimalMaintenanceHeuristic);
-    print_policy(d_optimalProductionHeuristic);
-    cout << heuristicCost << endl;
+    // Heuristic information
+
+    // Heuristic 1
+    outputFile.open("output/heuristic1_policy" + to_string(d_id) + ".csv");
+    num_t heuristicCost1 = cost_heuristic(true, false);
+    outputFile << print_policy_file(d_optimalMaintenanceHeuristic, d_optimalProductionHeuristic);
+    outputFile.close();
+
+    // Heuristic 2
+    outputFile.open("output/heuristic2_policy" + to_string(d_id) + ".csv");
+    num_t heuristicCost2 = cost_heuristic(false, true);
+    outputFile << print_policy_file(d_optimalMaintenanceHeuristic, d_optimalProductionHeuristic);
+    outputFile.close();
+
+    // Heuristic 3
+    outputFile.open("output/heuristic3_policy" + to_string(d_id) + ".csv");
+    num_t heuristicCost3 = cost_heuristic(true, true);
[…]

### 2018-05-31 src/run.cpp

@@ -20,12 +20,12 @@ try
         Init init{1,        // ccm
                   0.2,      // cpm
                   0.5,      // C
-                  20,       // nStates
+                  21,       // nStates
                   1,        // dL
-                  30,       // pi
-                  20,       // nRates
+                  20,       // pi
+                  21,       // nRates
                   1,        // mu
-                  2,        // var
+                  5,        // var
                   1,        // dT
                   0.1,      // beta
                   1.5,      // alpha
@@ -35,6 +35,7 @@ try
                  };
 
         // Initialise, solve and write mdp
+        init.printAllInfo();
         MDP mdp{init, false};
         mdp.solve();
         mdp.write_solution();

### 2018-06-01 report/paper_scm.tex
Updated report and experiments

@@
 For this reason, maintenance strategies have been extensively researched in the past decades.
-Reviews of past research have been published (CITE).
+Reviews of past research have been published \citep{VANHORENBEEK2013499, OLDEKEIZER2017405}.
 Condition-based maintenance (CBM) is an advanced maintenance strategy that is based on performance and/or parameter monitoring and basing subsequent actions on the monitoring.
@@
 \cite{OLDEKEIZER2018319} extend this research by including load sharing between units, that is, the deterioration of the functioning units is affected by the number of functioning units, where it is assumed that the load is shared uniformly over all functioning units.
-In this paper we consider a system where we relax this assumption and allow the load of all functioning units to depend on the current deterioration level of these units.
+In this paper, we consider a system where we relax this assumption and allow the load of all functioning units to depend on the current deterioration level of these units.
+Setting the load of the functioning units allows to cluster maintenance more effectively.
+In the rest of the paper, I will use the terminology production instead of load, however observe that production does not assumes the production of an item.
+As I will show in this paper, the effectiveness of setting the production (load) increases with the economic dependence between units.
 
 The coupling of production and maintenance has been the subject of several studies in recent years.
-Li and Ni (2009 Jin, X., L.
-Li, and J.
-Ni.
-2009.
-“Option Model for Joint Production and Preventive Maintenance System.” International Journal of Production Economics 119(2): 347–353.
-[Crossref], [Web of Science ®], [Google Scholar]
-) present an option-based cost model for scheduling a joint production and maintenance policy for a single-period, single-product system.
+\cite{doi:10.1080/00207543.2013.843037} present an option-based cost model for scheduling a joint production and maintenance policy for a single-period, single-product system.
+\cite{doi:10.1080/00207543.2016.1174343}  study an integrated control strategy of production and maintenance for a machining system which produces a single type of product to meet the constant demand.
+The authors consider a single-unit system with a preventive maintenance policy.
 
@@
 We formulate the system as a Markov Decision Process and analyse the structural properties of the optimal policies.
-Our model extends the model of \cite{OLDEKEIZER2018319} by allowing the load sharing to be set for all units.
+Our model extends the model of \cite{OLDEKEIZER2018319} by allowing the production (load) to be set for all units.
 The deterioration process of the units is modelled by a discrete-time Markov chain with a known transition probability matrix.
-The load sharing and maintenance strategy are jointly optimised, while in traditional methods the load sharing is set be uniform for all units.
+The production and maintenance strategy are jointly optimised, while in traditional methods the production is set be uniform for all functioning units.
 
@@
 Conclusions and future research directions are provided in Section 6.
-% For systems with multiple units, there exists dependence between the units.
-
-% Windmills require expensive fixed costs, so clustering maintenance can be profitable.
-In many windmill contracts a specified production has to be met for the total windpark.
-Try to combine production with maintenance policies.
 
 \section{System formulation}
-We consider a discrete-time system consisting of $N$ identical components, which are subject to load sharing and economic dependence.
+We consider a discrete-time system consisting of $N$ identical components, which are subject to variable production and economic dependence.
 We denote the set of units as $\mathcal{I} = \{1, \dots, n\}$.
-The total load that has to be met during every time period is denoted $\pi$.
+The total production that has to be met during every time period is denoted $\pi$.
 Let $x_i$ denote the deterioration state of unit $i$, which deteriorates according to a discrete-time Markov chain with deterioration states $X = \{0, 1, \dots, L\}$.
 State $0$ represents the as-good-as-new state, $L-1$ denotes the most deteriorated state such that the unit still functions and state $L$ the failed state.
-We assume that the load capacity of a unit is not affected by the deterioration state.
+We assume that the production capacity of a unit is not affected by the deterioration state.
 A schematic overview of the events in one time period is presented in Figure 1.
@@
 
-\subsection{Maintenance and load actions}
+\subsection{Maintenance and production actions}
 The deterioration state of each unit is known at the beginning of each time period.
@@
 As a performance criterion, we are interested in minimising the long-run average cost per time unit.
-In this way we can quantify the maintenance costs of a system over the life cycle, and identify the benefits of being able to set the load of each unit.
+In this way we can quantify the maintenance costs of a system over the life cycle, and identify the benefits of being able to set the production of each unit.
 
@@
 Similar to \cite{OLDEKEIZER2018319} we select preventive maintenance costs $c_\text{pm} = 5$, set-up cost $C = 4$, and a corrective maintenance cost $c_\text{cm} = 11$.
-The rate parameter under the maximum load rate is $\lambda_{\text{max}} = 0.7$ and the scale parameter $\gamma = 1$.
-The failure level $L$ and the number of possible production rates are set to 30.
-We vary the total load $\pi$ with values 30, 40 and 50.
-For $\pi = 50$, one unit can handle all the load and for $\pi = 50$ we do not have much overcapacity in the system.
+The rate parameter under the maximum production rate is $\lambda_{\text{max}} = 0.7$ and the scale parameter $\gamma = 1$.
+The failure level $L$ and the number of possible production rates are set to 25 and 26, respectively.
+We vary the total production $\pi$ with values 20, 30 and 40.
+For $\pi = 20$, one unit can handle all the production and for $\pi = 40$ we do not have much overcapacity in the system.
 We set $\alpha = 0.1$, such that the deterioration when a unit is not operating is low.
@@
 
-In the literature, condition-based maintenance is often implemented in the form of a threshold policy (CITE), or an opportunistic threshold policy \citep{ZHANG2015176}.
+In the literature, condition-based maintenance is often implemented in the form of a threshold policy (CITE), or an opportunistic threshold policy \citep{ZHANG2015176, ZHOU2009361} .
 In the former policy, if the deterioration level of a unit is at or exceed a threshold level $T_r$, we perform a maintenance action.
@@
 Due to the economic dependence (fixed cost of maintenance) and the liberty to set the production rate of the components, this policy is probably not optimal.
+From Figure ??
+observe that the optimal maintenance policy is not an opportunistic policy.
 
@@
 Additionally, we need a benchmark policy for the production policy.
-Inspired by the load sharing formulation in \cite{OLDEKEIZER2018319}, we will use the following benchmark for the production policy: For all system states determine the optimal selection of functioning units, where all functioning units share the load uniformly.
+Inspired by the load sharing formulation in \cite{OLDEKEIZER2018319}, we will use the following benchmark for the production policy: For all system states determine the optimal selection of functioning units, where the selected units share the total production $\pi$ uniformly.
 The optimal selection is found by minimising the long-run average  cost per time unit.
@@
 
-\subsection{Sensitivity analysis}
-\begin{enumerate}
-	\item Effect of fixed costs. More or less clustered maintenance.
-	\item Effect of variance.
+\subsection{Sensitivity analysis with respect to the maintenance set-up cost}
+To investigate how the optimal policy and relative effectiveness of the optimal policy compared to the heuristics varies for different levels of economic dependence
 
-\end{enumerate}
+\subsection{Sensitivity analysis with respect to variation in the deterioration process}
 
@@
 % Future research
-In this paper, I do not include dependence between the capacity of units and its deterioration state.
+Directions for future research in the direction of joint production and maintenance policies for multi-unit systems could be: 1) include dependence between the production capacity of units and its deterioration state, 2) include lead time for planning maintenance and 3) model dependence between the deterioration process of all units due to external shocks.
+The first extension could be realistic in some applications where the production of a unit is affected by the deterioration of that unit, e.g., semiconductor manufacturing \citep{kazaz2013impact}.
+The second direction is relevant for systems that have a high frequency of maintenance actions relative to the life cycle of the system.
+The last extension could be relevant when units are affected by, for example, weather conditions.
+It is realistic to assume that the deterioration of identical units is impacted in a similar way due to bad weather.
 

### 2018-06-01 create_experiments.R

@@ -5,16 +5,17 @@
 library(tidyverse)
 
 grid_size = 10
+problem_size = 11
 
 base_experiment = tibble(ccm = 11.0,
                          cpm = 5.0,
                          C = 4.0,
-                         nStates = 31,
+                         nStates = 26,
                          dL = 1.0,
-                         pi = 40,
-                         nRates = 31,
+                         pi = 48,
+                         nRates = 26,
                          mu = 1.0,
-                         var = 2.0,
+                         var = 1.0,
                          dT = 1.0,
                          minProb = 0.99,
                          beta = 0.1,
@@ -35,24 +36,24 @@ create_experiments = function(variableName, values) {
 }
 
 # Set pi
-pi_options = c(20, 30, 40, 58)
+pi_options = c(20, 30, 40)
 experiments = experiments %>%
   bind_rows(create_experiments("pi", pi_options))
 
 # Set alpha
-alpha_options = c(0, 1, 1.5)
+alpha_options = c(0, 0.5, 1)
 experiments = experiments %>%
   bind_rows(create_experiments("alpha", alpha_options))
 
-# # Variance experiments
-# var_options = seq(0.5, 5, length.out=grid_size)
-# experiments = experiments %>%
-#   bind_rows(create_experiments("var", var_options))
-# 
-# # Fixed cost effect
-# fixed_cost_options = seq(0, 3, length.out=grid_size)
-# experiments = experiments %>%
-#   bind_rows(create_experiments("C", fixed_cost_options))
+# Variance experiments
+var_options = seq(0.5, 5, length.out=grid_size)
+experiments = experiments %>%
+  bind_rows(create_experiments("var", var_options))
+
+# Fixed cost effect
+fixed_cost_options = seq(0, 3, length.out=grid_size)
+experiments = experiments %>%
+  bind_rows(create_experiments("C", fixed_cost_options))
 
 
 #### Output the experiments table ####

### 2018-06-01 create_figures.R

@@ -4,6 +4,7 @@
 
 library(tidyverse)
 library(reshape2)
+library(stringr)
 
 source("create_experiments.R")
 experiments = mutate(experiments, id = seq(0, nrow(experiments)-1))
@@ -30,10 +31,15 @@ idx = 0
 optimal_policy_data = NULL
 for (experiment_name in experiment_names) {
   optimal_policy = read_delim(paste(experiment_dir, experiment_name, sep="/"), delim=" ") %>%
-    mutate(`Maintenance action` = paste(maintenance_int_to_string(maintenance1), 
-                                        maintenance_int_to_string(maintenance2), 
-                                        sep=" ")) %>%
-    mutate(`Production action` = paste(production1, production2, sep=" ")) %>%
+    mutate(`Maintenance action` = case_when(
+      maintenance1 == 1 & maintenance2 == 1 ~ "Maintenance unit 1 and 2",
+      maintenance1 == 1 ~ "Maintenance unit 1",
+      maintenance2 == 1 ~ "Maintenance unit 2",
+      TRUE ~ "No maintenance")) %>%
+    mutate(`Production action` = case_when(
+      production1 == production2 ~ "equal",
+      production1 > production2 ~ "load1 > load2",
+      production2 > production1 ~ "load1 < load2")) %>%
     mutate(state1_max = state1 + 1,
            state2_max = state2 + 1,
            id = idx)
@@ -47,9 +53,6 @@ for (experiment_name in experiment_names) {
 # For each experiment we create the figures
 for (experiment_id in 0:(n_experiments-1)) {
   optimal_policy = filter(optimal_policy_data, id == experiment_id)
-  
-  # In wide format
-  optimal_policy_wide = dcast(optimal_policy, state1 ~ state2, value.var="Maintenance action")
   
   # Optimal maintenance with rectangles
   opt_maintenance_plot2 = ggplot(optimal_policy, aes(xmin=state1, xmax=state1_max,
@@ -66,7 +69,7 @@ for (experiment_id in 0:(n_experiments-1)) {
   # Plot optimal production with rectangles
   opt_production_plot2 = ggplot(optimal_policy, aes(xmin=state1, xmax=state1_max,
                                                     ymin=state2, ymax=state2_max,
-                                                    fill=production1)) +
+                                                    fill=`Production action`)) +
[…]

### 2018-06-01 src/mdp/feasible_production.cpp

@@ -6,7 +6,9 @@
 
 bool MDP::feasible_production(int state1, int state2, int prodRate1, int prodRate2) const
 {
-    if (prodRate1 + prodRate2 != d_pi)
+    if (prodRate1 < 0)
+        return false;
+    else if (prodRate1 + prodRate2 != d_pi)
         return false;
     
     return not (((state1 == (d_nStates - 1)) and (prodRate1 > 0)) or ((state2 == (d_nStates - 1)) and (prodRate2 > 0)));

### 2018-06-01 src/mdp/cost_heuristic.cpp
Fixed bug that last instead of optimal maintenance heuristic would show up

@@ -22,7 +22,10 @@ num_t MDP::cost_heuristic(bool mHeuristic, bool pHeuristic)
                 num_t const tmpCosts = bs::approximate_gain(d_valueBHeuristic, d_valueAHeuristic);
 
                 if (tmpCosts < minCosts)
+                {
                     minCosts = tmpCosts;
+                    d_optimalMaintenanceHeuristic = d_maintenanceHeuristic;
+                }
             }
         }
     } else 

### 2018-06-01 src/mdp/mdp.cpp

@@ -25,6 +25,7 @@ MDP::MDP(Init const &init, bool jumpHeuristic)
     d_eps(init.d_eps),
     d_optimalMaintenanceHeuristic(vector<array<int, 2>>(init.d_nStates * init.d_nStates, array<int, 2>())),
     d_optimalProductionHeuristic(vector<array<int, 2>>(init.d_nStates * init.d_nStates, array<int, 2>())),
+    d_maintenanceHeuristic(vector<array<int, 2>>(init.d_nStates * init.d_nStates, array<int, 2>())),
     d_valueAHeuristic(vector<num_t>(init.d_nStates * init.d_nStates, 0)),
     d_valueBHeuristic(vector<num_t>(init.d_nStates * init.d_nStates, 0)),
     d_expectedCostHeuristic(vector<num_t>(init.d_nStates * init.d_nStates, 0))

### 2018-06-01 src/mdp/mdp.h

@@ -54,6 +54,7 @@ class MDP
     // Heuristic
     std::vector<std::array<int, 2> > d_optimalMaintenanceHeuristic;
     std::vector<std::array<int, 2> > d_optimalProductionHeuristic;
+    std::vector<std::array<int, 2> > d_maintenanceHeuristic;
 
     std::vector<num_t> d_valueAHeuristic;
     std::vector<num_t> d_valueBHeuristic;

### 2018-06-01 src/mdp/set_maintenance_matrix.cpp

@@ -12,29 +12,29 @@ void MDP::set_maintenance_matrix(int T_r, int T_o)
         {
             if (sdx1 >= T_r)
             {
-                d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][0] = 1;
+                d_maintenanceHeuristic[hash(sdx1, sdx2)][0] = 1;
                 if (sdx2 >= T_o)
-                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][1] = 1;
+                    d_maintenanceHeuristic[hash(sdx1, sdx2)][1] = 1;
                 else
-                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][1] = 0;
+                    d_maintenanceHeuristic[hash(sdx1, sdx2)][1] = 0;
             } else if (sdx1 >= T_o)
             {
                 if (sdx2 >= T_r)
                 {
-                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][0] = 1;
-                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][1] = 1;
+                    d_maintenanceHeuristic[hash(sdx1, sdx2)][0] = 1;
+                    d_maintenanceHeuristic[hash(sdx1, sdx2)][1] = 1;
                 } else 
                 {
-                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][0] = 0;
-                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][1] = 0;
+                    d_maintenanceHeuristic[hash(sdx1, sdx2)][0] = 0;
+                    d_maintenanceHeuristic[hash(sdx1, sdx2)][1] = 0;
                 }
             } else 
             {
-                d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][0] = 0;
+                d_maintenanceHeuristic[hash(sdx1, sdx2)][0] = 0;
                 if (sdx2 >= T_r)
-                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][1] = 1;
+                    d_maintenanceHeuristic[hash(sdx1, sdx2)][1] = 1;
                 else
-                    d_optimalMaintenanceHeuristic[hash(sdx1, sdx2)][1] = 0;
+                    d_maintenanceHeuristic[hash(sdx1, sdx2)][1] = 0;
             }
         }
     }

### 2018-06-01 src/mdp/value_costs_heuristic.cpp

@@ -5,8 +5,8 @@
  */
 num_t MDP::value_costs_heuristic1(int state1, int state2, int idxState)
 {
-    int adx1 = d_optimalMaintenanceHeuristic[idxState][0];
-    int adx2 = d_optimalMaintenanceHeuristic[idxState][1];
+    int adx1 = d_maintenanceHeuristic[idxState][0];
+    int adx2 = d_maintenanceHeuristic[idxState][1];
     int s1 = (adx1 == 0) ? state1 : 0;
     int s2 = (adx2 == 0) ? state2 : 0;
 
@@ -54,8 +54,8 @@ num_t MDP::value_costs_heuristic2(int state1, int state2, int idxState)
  */
 num_t MDP::value_costs_heuristic3(int state1, int state2, int idxState)
 {
-    int adx1 = d_optimalMaintenanceHeuristic[idxState][0];
-    int adx2 = d_optimalMaintenanceHeuristic[idxState][1];
+    int adx1 = d_maintenanceHeuristic[idxState][0];
+    int adx2 = d_maintenanceHeuristic[idxState][1];
     int s1 = (adx1 == 0) ? state1 : 0;
     int s2 = (adx2 == 0) ? state2 : 0;
 

### 2018-06-01 report/paper_scm.tex
update report

@@
 \usepackage{dsfont}
+\usepackage{caption}
 
@@
 
+\textbf{[[Next paragraph is still detached from the rest of the story. Have to fit this in somehow. ]]}
 Offshore wind farms is one of the renewable energy solutions.
@@
 $$
-We assume that $0 \leq \beta \leq 1$ and $\alpha \geq 0$, since $g$ should be non-decreasing in $u_i$.
+We assume that $0 \leq \beta \leq 1$ and $\alpha \geq 0$, since $g$ should be non-decreasing in $u_i$ and positive for all production rates.
 If the production rate $u_i$ is zero, then $g$ has value $\beta$ i.e., $\beta$ determines the deterioration rate if the unit does not produce.
@@
 \bar{c}(x,r) = \begin{cases}
-C + \sum_{i=1}^{n}c(x_i, r_i), &\text{if } r \text{ contains a maintenance action},\\
-0, &\text{otherwise}.
+C + \sum_{i=1}^{n}c(x_i, r_i) &\text{if } r \text{ contains a maintenance action},\\
+0 &\text{otherwise}.
 \end{cases}
@@
 \section{Numerical results two-unit system}
-In this Section we will investigate the key features of the optimal maintenance and production policy for a two-unit system using the MDP as described in Section 3.
+In this section I will discuss the key features of the optimal maintenance and production policy for a two-unit system using the MDP as described in Section 3.
 Similar to \cite{OLDEKEIZER2018319} we select preventive maintenance costs $c_\text{pm} = 5$, set-up cost $C = 4$, and a corrective maintenance cost $c_\text{cm} = 11$.
@@
 The failure level $L$ and the number of possible production rates are set to 25 and 26, respectively.
-We vary the total production $\pi$ with values 20, 30 and 40.
-For $\pi = 20$, one unit can handle all the production and for $\pi = 40$ we do not have much overcapacity in the system.
-We set $\alpha = 0.1$, such that the deterioration when a unit is not operating is low.
+We vary the total production $\pi$ with values 20, 30, 40 and 48.
+For $\pi = 20$, one unit can handle all the production, whereas for $\pi = 48$ we do not have much overcapacity in the system.
+We set $\alpha = 0.1$, such that the deterioration rate is low when the unit is not producing.
 We consider three different values for $\beta$:  $0$ (deterioration rate does not depend on production), $1$ (deterioration rate depends linearly on production) and $1.5$ (deterioration rate is convex in the production).
+
+Figures 3-6 show the optimal policies systems with total production rates of 20, 30, 40 and 48.
+For every time period we first decide on the maintenance action and subsequently set a production rate for both units.
+For this reason we do not have to consider the production policies for states in which we perform maintenance.
+When the system can meet demand with one unit (Figure 3) we observe that maintenance is only performed if either both machines are in a highly deteriorated state, or one unit is a highly deteriorated state and the other is not.
+The last phenomenon could be explained by the fact that we want to bring the highly deteriorated unit back to the as-good-as-new state, such that, by setting its production rate higher than the other unit, we can bring the deterioration states close to each other to end up clustering the maintenance of both units.
+Furthermore, we observe that it is optimal to bring the deterioration state of the units close to each other by adjusting the production rate accordingly.
+
+As we increase the total production, we observe a different optimal production policy.
+When unit 1 (2) is in a `medium' deterioration state and unit 2 (1) is slightly higher (stated marked by the blue (pink) colour), it is optimal to try to bring the deterioration states close to each other by setting the production rate of unit 1 higher.
+However, when the difference in deterioration in sufficiently large, it is optimal to keep the expected difference in the deterioration state the same.
+\begin{figure}[H] \label{main-prod-policy}
+	\centering
+	\includegraphics[width=0.6\textwidth]{figures/optimal_action_plot1}
+	\caption{Optimal maintenance and production policy for a system with $N=2, L=25, m=25, c_\text{pm} = 5, c_\text{cm} = 11, C=4, \beta=0.1, \alpha=1.5, \pi=20$}
+\end{figure}
+\begin{figure}[H] \label{main-prod-policy2}
+	\centering
+	\includegraphics[width=0.6\textwidth]{figures/optimal_action_plot2}
+	\caption{Optimal maintenance and production policy for a system with $N=2, L=25, m=25, c_\text{pm} = 5, c_\text{cm} = 11, C=4, \beta=0.1, \alpha=1.5, \pi=30$}
+\end{figure}
+\begin{figure}[H] \label{main-prod-policy3}
+	\centering
+	\includegraphics[width=0.6\textwidth]{figures/optimal_action_plot3}
+	\caption{Optimal maintenance and production policy for a system with $N=2, L=25, m=25, c_\text{pm} = 5, c_\text{cm} = 11, C=4, \beta=0.1, \alpha=1.5, \pi=40$}
+\end{figure}
+\begin{figure}[H] \label{main-prod-policy4}
+	\centering
+	\includegraphics[width=0.6\textwidth]{figures/optimal_action_plot0}
+	\caption{Optimal maintenance and production policy for a system with $N=2, L=25, m=25, c_\text{pm} = 5, c_\text{cm} = 11, C=4, \beta=0.1, \alpha=1.5, \pi=48$}
+\end{figure}
 
@@
 
-Table (???) presents the
+To compare the optimal policy of the maintenance and production policy to the heuristic algorithms, we define the following combinations of heuristics:
+\begin{itemize}
+	\item \textit{Heuristic 1}: maintenance policy determined by the heuristic algorithm and an optimal production policy;
+	\item \textit{Heuristic 2}: an optimal maintenance policy and a heuristic production policy;
+	\item \textit{Heuristic 3}: heuristics for both the maintenance and production policy.
+\end{itemize}
+Heuristic 2 is used to quantifies the benefit of allowing for a variable production rate, whereas heuristic 3 is combination of popular policies that are widely used in practice.
+
+Table 1 shows that for the different total productions the additional average costs can be as much as $3.5$\%.
+The additional average costs for heuristic 1 is lower than heuristic 2, \textbf{[[So....]]}.
+
+\ctable[
+pos=h,
+caption={The minimal long-run average cost per time unit for the heuristic policies with different values for $\pi$}
+]{l c c c c}{}{
+	\FL
+	Total production & \multicolumn{3}{c}{Average cost} & Optimal thresholds \NN
+	 & Heuristic 1 & Heuristic 2 & Heuristic 3 & Heuristic 1 \ML
+	$\pi = 20$ & 0.1307 (+0.06\%) & 0.1320 (+1.80\%) & 0.1351 (+3.46\%) & $T_r = 24$, $T_o = 16$ \NN
+	$\pi = 30$ & 0.6378 (+0.34\%) & 0.6404 (+0.76\%) & 0.6417 (+0.96\%) & $T_r = 23$, $T_o = 15$ \NN
+	$\pi = 40$ & 0.6521 (+0.19\%) & 0.6543 (+0.56\%) & 0.6552 (+0.66\%) & $T_r = 23$, $T_o = 13$ \NN
+	$\pi = 48$ & 0.5976 (+0.11\%) & 0.6047 (+1.28\%)& 0.6052 (+1.38\%) & $T_r = 23$, $T_o = 11$ \LL
+}
 
@@
 
+\section{Notes}
+\begin{itemize}
+	\item I have not figured out yet how to set figures next to each other. My apologies. 
+	\item The colours of the optimal policy plots are disturbing. I do have ideas for different visualisations, however if you have any ideas, feel free to write them in your review.
+	\item The practical relevance of heuristic 1 is still unclear for me. I might leave it out.
+	\item I do not show absolute values of production in the optimal policy plot to not clutter the message, however I feel like key features of the optimal policy are missing in this way.
+\end{itemize}
+
 \bibliographystyle{IEEEtranSN} 

### 2018-06-01 create_figures.R

@@ -37,12 +37,14 @@ for (experiment_name in experiment_names) {
       maintenance2 == 1 ~ "Maintenance unit 2",
       TRUE ~ "No maintenance")) %>%
     mutate(`Production action` = case_when(
-      production1 == production2 ~ "equal",
-      production1 > production2 ~ "load1 > load2",
-      production2 > production1 ~ "load1 < load2")) %>%
+      production1 == production2 ~ "equal production",
+      production1 > production2 ~ "production1 > production2",
+      production2 > production1 ~ "production1 < production2")) %>%
     mutate(state1_max = state1 + 1,
            state2_max = state2 + 1,
-           id = idx)
+           id = idx) %>%
+    mutate(action = if_else(`Maintenance action` == "No maintenance", 
+                            `Production action`, `Maintenance action`))
   
   optimal_policy_data = optimal_policy_data %>%
     bind_rows(optimal_policy)
@@ -78,28 +80,46 @@ for (experiment_id in 0:(n_experiments-1)) {
   ggsave(paste("output/optimal_production_plot", experiment_id, ".pdf", sep=""), 
          opt_production_plot2, width=9, height=7)
   
-  # Combine both plots
+  # Combine plot
+  opt_action_plot = ggplot(optimal_policy, aes(xmin=state1, xmax=state1_max,
+                                               ymin=state2, ymax=state2_max,
+                                               fill=action)) + 
+    geom_rect() + scale_colour_grey() + theme_bw() + theme(panel.grid = element_blank(), panel.border = element_blank()) +
+    scale_x_continuous(expand = c(0, 0)) +
+    scale_y_continuous(expand = c(0, 0)) +
+    labs(x="State unit 1", y="State unit 2") + 
+  
+  ggsave(paste("output/optimal_action_plot", experiment_id, ".pdf", sep=""),
+         width=9, height=7)
+  
 }
 
 
 #### Average gain plots #####
 
 avg_gain_names = list.files(experiment_dir, pattern = "^approximate_gain[0-9]*$")
+avg_gain_heuristic_names = list.files(experiment_dir, pattern = "^approximate_gain_heuristics[0-9]*$")
 n_experiments = length(avg_gain_names)-1
 approx_gains = NULL
 
+# Load optimal values
 for (experiment_id in 0:n_experiments) {
[…]

### 2018-06-01 src/mdp/write_solution.cpp

@@ -20,7 +20,6 @@ void MDP::write_solution()
     outputFile.open("output/optimal_values" + to_string(d_id) + ".csv");
     for (num_t const val: d_valueB)
         outputFile << val << ' ';
-    
     outputFile.close();
 
     // Save approximation gain
@@ -31,20 +30,20 @@ void MDP::write_solution()
     // Heuristic information
 
     // Heuristic 1
-    outputFile.open("output/heuristic1_policy" + to_string(d_id) + ".csv");
     num_t heuristicCost1 = cost_heuristic(true, false);
+    outputFile.open("output/heuristic1_policy" + to_string(d_id) + ".csv");
     outputFile << print_policy_file(d_optimalMaintenanceHeuristic, d_optimalProductionHeuristic);
     outputFile.close();
 
     // Heuristic 2
-    outputFile.open("output/heuristic2_policy" + to_string(d_id) + ".csv");
     num_t heuristicCost2 = cost_heuristic(false, true);
+    outputFile.open("output/heuristic2_policy" + to_string(d_id) + ".csv");
     outputFile << print_policy_file(d_optimalMaintenanceHeuristic, d_optimalProductionHeuristic);
     outputFile.close();
 
     // Heuristic 3
-    outputFile.open("output/heuristic3_policy" + to_string(d_id) + ".csv");
     num_t heuristicCost3 = cost_heuristic(true, true);
+    outputFile.open("output/heuristic3_policy" + to_string(d_id) + ".csv");
     outputFile << print_policy_file(d_optimalMaintenanceHeuristic, d_optimalProductionHeuristic);
     outputFile.close();
 

### 2018-06-05 report/paper_scm.tex
Added function to write heuristic values to table

@@
 
+The coupling of production and maintenance for single-unit systems has been the subject of several studies in recent years.
+\cite{doi:10.1080/00207543.2013.843037} present an option-based cost model for scheduling a joint production and maintenance policy for a single-period, single-product system.
+\cite{doi:10.1080/00207543.2016.1174343}  study an integrated control strategy of production and maintenance for a machining system which produces a single type of product to meet the constant demand.
+The authors consider a single-unit system with a preventive maintenance policy.
+
 Condition-based maintenance policies have been proposed for multi-unit systems with economic dependence \citep{OLDEKEIZER2016531}.
 \cite{OLDEKEIZER2018319} extend this research by including load sharing between units, that is, the deterioration of the functioning units is affected by the number of functioning units, where it is assumed that the load is shared uniformly over all functioning units.
-In this paper, we consider a system where we relax this assumption and allow the load of all functioning units to depend on the current deterioration level of these units.
+In this paper, we consider a system where we generalise this model and allow the load of all functioning units to depend on the current deterioration level of these units.
 Setting the load of the functioning units allows to cluster maintenance more effectively.
@@
 As I will show in this paper, the effectiveness of setting the production (load) increases with the economic dependence between units.
-
-The coupling of production and maintenance has been the subject of several studies in recent years.
-\cite{doi:10.1080/00207543.2013.843037} present an option-based cost model for scheduling a joint production and maintenance policy for a single-period, single-product system.
-\cite{doi:10.1080/00207543.2016.1174343}  study an integrated control strategy of production and maintenance for a machining system which produces a single type of product to meet the constant demand.
-The authors consider a single-unit system with a preventive maintenance policy.
 
@@
 \cite{krokoszinski2003efficiency}.
-Existing CBM methods for offshore wind farm systems deal with wind turbines as individual components separately \citep{avner2007reliability}.
+Existing CBM methods for offshore wind farm systems deal with wind turbines as individual units separately \citep{avner2007reliability}.
 Later maintenance strategies where economic dependence for maintenance actions between wind turbines is taken into account have gained interest \citep{tian2011condition}.
@@
 This reduction is mostly due to the fact that it exploits the large fixed costs from and cluster maintenance actions.
-One element of the fixed costs component is the cost of sending a maintenance team to the wind farm,, which generally requires expensive vessels.
+One element of the fixed costs unit is the cost of sending a maintenance team to the wind farm,, which generally requires expensive vessels.
 
-In this paper, we consider a joint production and maintenance strategy for a multi-unit system.
-We formulate the system as a Markov Decision Process and analyse the structural properties of the optimal policies.
-Our model extends the model of \cite{OLDEKEIZER2018319} by allowing the production (load) to be set for all units.
+In this paper, I consider a joint production and maintenance strategy for a multi-unit system.
+I formulate the system as a Markov Decision Process and analyse the structural properties of the optimal policies.
+The model extends the model of \cite{OLDEKEIZER2018319} by allowing the production (load) to be set for all units.
 The deterioration process of the units is modelled by a discrete-time Markov chain with a known transition probability matrix.
@@
 \section{System formulation}
-We consider a discrete-time system consisting of $N$ identical components, which are subject to variable production and economic dependence.
+We consider a discrete-time system consisting of $N$ identical units, which are subject to variable production and economic dependence.
 We denote the set of units as $\mathcal{I} = \{1, \dots, n\}$.
-The total production that has to be met during every time period is denoted $\pi$.
+The total production rate that has to be met during every time period is denoted $\pi$.
 Let $x_i$ denote the deterioration state of unit $i$, which deteriorates according to a discrete-time Markov chain with deterioration states $X = \{0, 1, \dots, L\}$.
@@
 Generally, the cost of preventive maintenance is lower than the corrective maintenance action ($c_\text{pm} < c_\text{cm}$).
-Performing maintenance on a unit instantanously brings it back to the as-good-as new state.
+Performing maintenance on a unit instantaneously brings it back to the as-good-as new state.
 The economic dependence is included in the form of a fixed cost $C$ for performing maintenance for a time period.
@@
 Each unit $i$ deteriorates according to a gamma process with rate $\lambda(u_i)$, which depends on the current production rate of the unit, and scale parameter $\gamma$.
+The gamma process is a rather flexible process that is applicable to model a wide variety of applications.
+According to \cite{van2009survey}, the gamma process is most appropriate to model deterioration processes.
+
 The rate for a given production rate $u_i$ is given by
@@
 \subsection{Action space}
-At the start of each time unit we have to decide on which components to maintain (M) and do nothing (DN).
+At the start of each time unit we have to decide on which units to maintain (M) and do nothing (DN).
 Maintenance can be performed one all units, independent of the current state, i.e., the action space for the maintenance decision can be defined as
@@
 Let $u = (u_1, \dots, u_n)$ denote the production rate of all units in the system.
-The total production of the system is restricted to $\pi$, i.e., $ \sum_{i=1}^{n}u_i = \pi$.
+The total production rate of the system is restricted to $\pi$, i.e., $ \sum_{i=1}^{n}u_i = \pi$.
 This results in the following action space for the system in state $x$
@@
 
-In the literature, condition-based maintenance is often implemented in the form of a threshold policy (CITE), or an opportunistic threshold policy \citep{ZHANG2015176, ZHOU2009361} .
-In the former policy, if the deterioration level of a unit is at or exceed a threshold level $T_r$, we perform a maintenance action.
-The opportunistic policy adds an additional step to the threshold policy: if an maintenance action is initiated, then all units that have a deterioration state higher or equal to $T_o$ are also maintained in that time period.
-The optimal threshold values for $T_r$ and $T_o$ are found using a grid search procedure.
-Due to the economic dependence (fixed cost of maintenance) and the liberty to set the production rate of the components, this policy is probably not optimal.
-From Figure ??
-observe that the optimal maintenance policy is not an opportunistic policy.
-
-\textbf{[[Note: I think it is unfair to compare the optimal policy to the standard threshold policy, since this policy is equal or worse than the opportunistic policy. ]]}
+In the literature, condition-based maintenance is often implemented in the form of an opportunistic threshold policy \citep{ZHANG2015176, ZHOU2009361}.
+For each time period $t$ this heuristic works as follows:
+\begin{enumerate}[1)]
+	\item Observe the deterioration states of all units, $i = 1,\dots, n$.
+	\item Maintain all units which have a deterioration state that exceed or equal a threshold $T_r$, i.e., if $X_i(t) \geq T_r$, $i = 1, \dots, n$.
+	\item If at least one maintenance action is initiated in step 2), also maintain all units for which $X_i(t) \geq T_o$, $i = 1, \dots, n$,
+\end{enumerate}
+where $T_o \leq T_r$.
+The optimal threshold values for $T_r$ and $T_o$ are found using a grid search procedure, where the objective is to minimise the long-run average cost.
+Due to the economic dependence (fixed cost of maintenance) and the ability to set the production rate of the units, this policy is probably not optimal.
+From Figure 3, observe that the optimal maintenance policy is not an opportunistic policy.
 

### 2018-06-05 src/mdp/mdp.h

@@ -72,6 +72,8 @@ class MDP
         void        print_policy(std::vector<std::array<int, 2>> const &policy) const;
         std::string print_policy_file(std::vector<std::array<int, 2>> const &maintenancePolicy,
                                       std::vector<std::array<int, 2>> const &productionPolicy) const;
+        void        print_latex_table(num_t optimalCost, num_t heuristicCost1, num_t heuristicCost2,
+                                      num_t heuristicCost3) const;
         
         size_t get_id() const;
 

### 2018-06-05 src/mdp/print_latex_table.cpp

@@ -1,0 +1,22 @@
+#include "mdp.ih"
+
+/*
+ * Create row for optimal cost table.
+ */
+
+void MDP::print_latex_table(num_t optimalCost, num_t heuristicCost1, num_t heuristicCost2,
+                            num_t heuristicCost3) const
+{
+    ofstream outputFile;
+
+    // If it is first experiment, clear file, otherwise append file.
+    if (d_id == 0)
+        outputFile.open("output/cost_table.tex");
+    else
+        outputFile.open("output/cost_table.tex", ofstream::app);
+
+    outputFile << heuristicCost1 << " (" << ((heuristicCost1 - optimalCost) / optimalCost) * 100 << "%)" << "&";
+    outputFile << heuristicCost2 << " (" << ((heuristicCost2 - optimalCost) / optimalCost) * 100 << "%)" << "&";
+    outputFile << heuristicCost3 << " (" << ((heuristicCost3 - optimalCost) / optimalCost) * 100 << "%)" << "//" << endl;
+    outputFile.close();
+}

### 2018-06-05 src/mdp/write_solution.cpp

@@ -23,8 +23,9 @@ void MDP::write_solution()
     outputFile.close();
 
     // Save approximation gain
+    num_t optimalCost = bs::approximate_gain(d_valueB, d_valueA);
     outputFile.open("output/approximate_gain" + to_string(d_id));
-    outputFile << bs::approximate_gain(d_valueB, d_valueA) << endl;
+    outputFile << optimalCost << endl;
     outputFile.close();
 
     // Heuristic information

### 2018-06-20 report/paper_scm.tex
Rewrote introduction and appended system formulation

@@
 
-\title{Maintenance Policies for a Multi-Unit System with Adjustable Production Rates}
+\title{Condition-based Maintenance and Production Policies for Multi-Unit Systems}
 \author{Mark van der Broek}
@@
 % Good maintenance policies are required
-Maintenance cost can constitute a large part of the total production costs.
-For example, \cite{GRABER2004861} states that maintenance costs can be at much as 30 percent of production costs for power plants, while \cite{BEVILACQUA200071} give an estimate of between 15 and 70 percent of the total production costs.
-For this reason, maintenance strategies have been extensively researched in the past decades.
-Reviews of past research have been published \citep{VANHORENBEEK2013499, OLDEKEIZER2017405}.
-Condition-based maintenance (CBM) is an advanced maintenance strategy that is based on performance and/or parameter monitoring and basing subsequent actions on the monitoring.
+A large part of the total production costs is due to maintenance.
+\cite{GRABER2004861} states that maintenance costs can be at much as 30 percent of production costs for power plants, while \cite{BEVILACQUA200071} give an estimate of between 15 and 70 percent of the total production costs.
+\cite{wang2002survey} and \cite{ding2015maintenance} give a review of past research on maintenance policies.
+An increasingly popular maintenance strategy is condition-based maintenance (CBM), which uses conditioning information about components of a system to designs maintenance policies.
+Despite the increasing popularity of CBM, the implementation of CBM policies in practical applications is lagging behind \citep{keizer2017condition}.
+For this reason, research on CBM involves more complex systems with multiple dependent components \citep{}.
+
+A general type of system is the $k$-out-of-$N$ system, which consists of $N$ components, and is functioning when at least $k$ components are functioning.
+\cite{OLDEKEIZER2018319} consider a $1$-out-of-$N$ system with economic dependence by imposing a fixed cost for performing maintenance and failure dependence by load sharing between units, that is, the deterioration of the functioning units is affected by the number of functioning units.
+The load of the functioning units is assumed to be shared uniformly over all functioning units.
+
+In this paper, I consider a $k$-out-of-$N$ system with load sharing, which allows the load of all functioning units to depend on the current deterioration level of these units.
+Setting the load of the functioning units allows to cluster maintenance more effectively, due to greater control of the deterioration of the components of the total system.
+In the rest of the paper, I will use the terminology production instead of load, however observe that production does not assumes the production of an item.
+As I will show in this paper, the effectiveness of setting the production (load) increases with the economic dependence between units.
 
@@
 \cite{doi:10.1080/00207543.2013.843037} present an option-based cost model for scheduling a joint production and maintenance policy for a single-period, single-product system.
-\cite{doi:10.1080/00207543.2016.1174343}  study an integrated control strategy of production and maintenance for a machining system which produces a single type of product to meet the constant demand.
+\cite{doi:10.1080/00207543.2016.1174343} study an integrated control strategy of production and maintenance for a machining system which produces a single type of product to meet the deterministic demand.
 The authors consider a single-unit system with a preventive maintenance policy.
-
-Condition-based maintenance policies have been proposed for multi-unit systems with economic dependence \citep{OLDEKEIZER2016531}.
-\cite{OLDEKEIZER2018319} extend this research by including load sharing between units, that is, the deterioration of the functioning units is affected by the number of functioning units, where it is assumed that the load is shared uniformly over all functioning units.
-In this paper, we consider a system where we generalise this model and allow the load of all functioning units to depend on the current deterioration level of these units.
-Setting the load of the functioning units allows to cluster maintenance more effectively.
-In the rest of the paper, I will use the terminology production instead of load, however observe that production does not assumes the production of an item.
-As I will show in this paper, the effectiveness of setting the production (load) increases with the economic dependence between units.
 
@@
 The deterioration process of the units is modelled by a discrete-time Markov chain with a known transition probability matrix.
-The production and maintenance strategy are jointly optimised, while in traditional methods the production is set be uniform for all functioning units.
+The production and maintenance strategy are jointly optimised, while $k$-out-of-$N$ models usually assume that the production is shared uniformly over all functioning units.
 
@@
 For all other deteriorating levels, we can select one of the $m+1$ production rates.
+
+Since I assume that we can perform maintenance instantaneously, contrary to the work of \cite{OLDEKEIZER2018319}, I do impose a penalty if the system does not function, i.e., the number of functioning units is less than $k$.
+Instead, I restrict the set of feasible maintenance policies to be such that we need at least $k$ functioning units after the maintenance action.
 
@@
 \subsection{Performance criteria}
-As a performance criterion, we are interested in minimising the long-run average cost per time unit.
-In this way we can quantify the maintenance costs of a system over the life cycle, and identify the benefits of being able to set the production of each unit.
+The performance criteria is to minimise the long-run average (maintenance) cost per time unit.
+This way, we can compare different condition-based maintenance and production strategies over the life cycle of the system.
 

### 2018-06-20 create_figures.R

@@ -186,7 +186,7 @@ for (heuristic_id in 1:3) {
            opt_maintenance_plot2, width=9, height=7)
     
     # Plot optimal production with rectangles
-    opt_production_plot2 = ggplot(optimal_policy, aes(xmin=state1, xmax=state1_max,
+    opt_production_plot2 = ggplot(heuristic_policy, aes(xmin=state1, xmax=state1_max,
                                                       ymin=state2, ymax=state2_max,
                                                       fill=`Production action`)) +
       geom_rect() + theme_bw() + theme(panel.grid = element_blank(), panel.border = element_blank()) +
@@ -199,7 +199,7 @@ for (heuristic_id in 1:3) {
     
     # Combine both plots
     # Combine plot
-    opt_action_plot = ggplot(optimal_policy, aes(xmin=state1, xmax=state1_max,
+    opt_action_plot = ggplot(heuristic_policy, aes(xmin=state1, xmax=state1_max,
                                                  ymin=state2, ymax=state2_max,
                                                  fill=action)) + 
       geom_rect() + scale_colour_grey() + theme_bw() + theme(panel.grid = element_blank(), panel.border = element_blank()) +

### 2018-06-20 src/mdp/expected_cost_heuristic.cpp

@@ -13,13 +13,16 @@ num_t MDP::expected_cost_heuristic1(int state1, int state2, int idxState)
     // Check if we should do one working unit
     if (d_pi <= d_nRates)
     {
-        bestVal = expected_costs_heuristic(state1, state2, d_pi, 0);
-        bestProd1 = d_pi;
-        bestProd2 = 0;
+        if (state1 != (d_nStates-1))
+        {
+            bestVal = expected_costs_heuristic(state1, state2, d_pi, 0);
+            bestProd1 = d_pi;
+            bestProd2 = 0;
+        }
 
         num_t const tmpVal2 = expected_costs_heuristic(state1, state2, 0, d_pi);
 
-        if (tmpVal2 < bestVal)
+        if (tmpVal2 < bestVal and state2 != (d_nStates-1))
         {
             bestVal = tmpVal2;
             bestProd1 = 0;
@@ -32,7 +35,7 @@ num_t MDP::expected_cost_heuristic1(int state1, int state2, int idxState)
     int pdx2 = d_pi - pdx1;
     num_t tmpVal3 = expected_costs_heuristic(state1, state2, pdx1, pdx2);
 
-    if (tmpVal3 < bestVal)
+    if (tmpVal3 < bestVal and feasible_production(state1, state2, pdx1, pdx2))
     {
         bestVal = tmpVal3;
         bestProd1 = pdx1;

### 2018-06-20 src/mdp/value_costs.cpp

@@ -22,7 +22,7 @@ num_t MDP::value_costs(int state1, int state2, int idxState)
             int s2 = (adx2 == 0) ? state2 : 0;
 
             num_t const tmpVal = direct_cost(state1, state2, adx1, adx2) + d_expectedCost[hash(s1, s2)];
-            if (tmpVal < bestVal - gg_eps)
+            if (tmpVal < bestVal)
             {
                 bestVal = tmpVal;
                 bestMain1 = adx1;

### 2018-06-20 src/mdp/value_costs_heuristic.cpp

@@ -33,7 +33,7 @@ num_t MDP::value_costs_heuristic2(int state1, int state2, int idxState)
             int s2 = (adx2 == 0) ? state2 : 0;
 
             num_t const tmpVal = direct_cost(state1, state2, adx1, adx2) + d_expectedCostHeuristic[hash(s1, s2)];
-            if (tmpVal < bestVal - gg_eps)
+            if (tmpVal < bestVal)
             {
                 bestVal = tmpVal;
                 bestMain1 = adx1;

### 2018-06-20 src/solve_experiments.h

@@ -10,8 +10,10 @@
 void solve_experiment(Init const &init)
 {
     // Initialise MDP
+    init.printAllInfo();
     MDP mdp{init, false};
     mdp.solve();
+    mdp.print_all_info();
     mdp.write_solution();
     std::cout << "Solved experiment " << mdp.get_id() << std::endl;
 }
@@ -29,7 +31,6 @@ void solve_from_file()
         while (getline(experimentsFile, line))
         {
             Init init{line};
-            init.printAllInfo();
             init_vec.push_back(init);
         }
         experimentsFile.close();

### 2018-06-21 src/mdp/expected_costs.cpp
Clean up code

@@ -16,11 +16,11 @@ num_t MDP::expected_costs(int state1, int state2, int prodRate1, int prodRate2) 
     int maxJump1 = d_nStates-state1-1;
     int maxJump2 = d_nStates-state2-1;
 
-    if (d_jumpHeuristic) 
-    {
-        maxJump1 = min(maxJump1, d_maxJumps[prodRate1]);
-        maxJump2 = min(maxJump2, d_maxJumps[prodRate2]);
-    } 
+    // if (d_jumpHeuristic) 
+    // {
+    //     maxJump1 = min(maxJump1, d_maxJumps[prodRate1]);
+    //     maxJump2 = min(maxJump2, d_maxJumps[prodRate2]);
+    // } 
 
     // Loop over all jump probabilities from 'state1' to 'd_nStates' and obtain costs.
     int idxState = hash(state1, 0);
@@ -39,9 +39,7 @@ num_t MDP::expected_costs(int state1, int state2, int prodRate1, int prodRate2) 
     idxState += state2;
 
     for (int sdx2 = 0; sdx2 != maxJump2; ++sdx2)
-    {
         costs += prop1 * d_prop1[prodRate2][sdx2] * d_valueA[idxState++];
-    }
     costs += prop1 * d_prop2[prodRate2][state2] * d_valueA[idxState++];
 
     return costs;

### 2018-06-22 report/paper_scm.tex
Final changes to model

@@
 $$
+
 Note that the production action space decreases from $(m+1)^n$ elements to ${n + \pi - 1} \choose{\pi - 1} $ by restriction the total production rate to the value $\pi$.
+Observe that a $1$-out-of-$N$ can be obtained if we set $\pi$ to 1, and, more generally, a $k$-out-of-$N$ system if we set $\pi$ equal to $k$.
 
@@
 In this section I will discuss the key features of the optimal maintenance and production policy for a two-unit system using the MDP as described in Section 3.
-Similar to \cite{OLDEKEIZER2018319} we select preventive maintenance costs $c_\text{pm} = 5$, set-up cost $C = 4$, and a corrective maintenance cost $c_\text{cm} = 11$.
-The rate parameter under the maximum production rate is $\lambda_{\text{max}} = 0.7$ and the scale parameter $\gamma = 1$.
-The failure level $L$ and the number of possible production rates are set to 25 and 26, respectively.
-We vary the total production $\pi$ with values 20, 30, 40 and 48.
-For $\pi = 20$, one unit can handle all the production, whereas for $\pi = 48$ we do not have much overcapacity in the system.
-We set $\alpha = 0.1$, such that the deterioration rate is low when the unit is not producing.
-We consider three different values for $\beta$:  $0$ (deterioration rate does not depend on production), $1$ (deterioration rate depends linearly on production) and $1.5$ (deterioration rate is convex in the production).
+I normalise the corrective maintenance cost to $c_\text{cm}$ to 1, the fixed cost is set to $C = 0.4$ and preventative maintenance $c_\text{pm}$ to 0.2, i.e., the cost of performing corrective maintenance is five times higher than preventive maintenance.
+For the deterioration process we set the mean and variance equal to 1.
+The rate and scale parameter of the gamma process are implicitly defined for any choice of the mean and variance parameter.
+The number of deterioration states and production actions for each unit is set to 26, i.e., $L = 25$ and $m = 25$.
+The base total production value $\pi$ is set to 1 (which results in a $1$-out-of-$N$ system).
+The parameters for the function $g$ are set to $\alpha = 0.1$, such that the deterioration rate is low when the unit is not producing, and $\beta = 1.5$.
 

### 2018-06-22 create_experiments.R

@@ -4,23 +4,23 @@
 
 library(tidyverse)
 
-grid_size = 10
-problem_size = 11
+grid_size = 30
+problem_size = 25
 
-base_experiment = tibble(ccm = 11.0,
-                         cpm = 5.0,
-                         C = 4.0,
-                         nStates = 26,
+base_experiment = tibble(ccm = 100,
+                         cpm = 0.2 * ccm,
+                         C = 0.4 * ccm,
+                         nStates = problem_size + 1,
                          dL = 1.0,
-                         pi = 48,
-                         nRates = 26,
+                         pi = problem_size,
+                         nRates = problem_size + 1,
                          mu = 1.0,
                          var = 1.0,
                          dT = 1.0,
                          minProb = 0.99,
                          beta = 0.1,
                          alpha = 1.5,
-                         eps = 0.000001,
+                         eps = 0.0000001,
                          nUnits = 2)
 
 experiments = base_experiment
@@ -35,8 +35,33 @@ create_experiments = function(variableName, values) {
   return(new_experiments)
 }
 
-# Set pi
-pi_options = c(20, 30, 40)
+# Variance experiments
+var_options = seq(0.5, 5, length.out=grid_size)
+experiments = experiments %>%
+  bind_rows(create_experiments("var", var_options))
+
+# Fixed cost effect
+fixed_cost_options = seq(0, 200, length.out=grid_size)
+experiments = experiments %>%
+  bind_rows(create_experiments("C", fixed_cost_options))
+
+# Preventive effect 
+preventive_cost_options = seq(0, 100, length.out=grid_size)
+experiments = experiments %>%
+  bind_rows(create_experiments("cpm", preventive_cost_options))
+
+# Beta
+b_options = seq(0, 1, length.out=grid_size)
+experiments = experiments %>%
+  bind_rows(create_experiments("beta", b_options))
+
+# Alpha
+alpha_options = seq(0, 3, length.out=grid_size)
+experiments = experiments %>%
+  bind_rows(create_experiments("alpha", alpha_options))
+
+# Pi
+pi_options = seq(20, 50, by=5)
 experiments = experiments %>%
   bind_rows(create_experiments("pi", pi_options))
 
@@ -44,16 +69,6 @@ experiments = experiments %>%
 alpha_options = c(0, 0.5, 1)
 experiments = experiments %>%
   bind_rows(create_experiments("alpha", alpha_options))
-
-# Variance experiments
[…]

### 2018-06-22 create_figures.R

@@ -5,6 +5,7 @@
 library(tidyverse)
 library(reshape2)
 library(stringr)
+library(ggthemes)
 
 source("create_experiments.R")
 experiments = mutate(experiments, id = seq(0, nrow(experiments)-1))
@@ -32,23 +33,24 @@ optimal_policy_data = NULL
 for (experiment_name in experiment_names) {
   optimal_policy = read_delim(paste(experiment_dir, experiment_name, sep="/"), delim=" ") %>%
     mutate(`Maintenance action` = case_when(
-      maintenance1 == 1 & maintenance2 == 1 ~ "Maintenance unit 1 and 2",
-      maintenance1 == 1 ~ "Maintenance unit 1",
-      maintenance2 == 1 ~ "Maintenance unit 2",
-      TRUE ~ "No maintenance")) %>%
+      maintenance1 == 1 & maintenance2 == 1 ~ "12",
+      maintenance1 == 1 ~ "1",
+      maintenance2 == 1 ~ "2",
+      TRUE ~ "MNA")) %>%
+    mutate(production1 = production1 / problem_size,
+           production2 = production2 / problem_size) %>%
     mutate(`Production action` = case_when(
-      production1 == production2 ~ "equal production",
-      production1 > production2 ~ "production1 > production2",
-      production2 > production1 ~ "production1 < production2")) %>%
+      production1 == production2 ~ "=",
+      production1 > production2 ~ "+",
+      production2 > production1 ~ "-")) %>%
     mutate(state1_max = state1 + 1,
            state2_max = state2 + 1,
-           id = idx) %>%
-    mutate(action = if_else(`Maintenance action` == "No maintenance", 
+           id = as.numeric(str_extract(experiment_name, "[0-9]+"))) %>%
+    mutate(action = if_else(`Maintenance action` == "MNA", 
                             `Production action`, `Maintenance action`))
   
   optimal_policy_data = optimal_policy_data %>%
     bind_rows(optimal_policy)
-  idx = idx + 1
 }
 
 
@@ -56,38 +58,31 @@ for (experiment_name in experiment_names) {
 for (experiment_id in 0:(n_experiments-1)) {
   optimal_policy = filter(optimal_policy_data, id == experiment_id)
   
-  # Optimal maintenance with rectangles
-  opt_maintenance_plot2 = ggplot(optimal_policy, aes(xmin=state1, xmax=state1_max,
[…]

### 2018-06-22 peregrind_job.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+
+#SBATCH --job-name=MDP
+#SBATCH --time=08:00:00
+#SBATCH --cpus-per-task=24
+#SBATCH --mem=1000
+
+# Load modules
+module load foss/2018a
+
+# Run the code
+make clean
+make
+./mdpsolver -f

### 2018-06-22 src/init/init.cpp

@@ -2,10 +2,13 @@
 
 #include <boost/algorithm/string.hpp>
 
+size_t Init::s_id = 0;
+
 Init::Init(num_t ccm, num_t cpm, num_t C, int nStates, num_t dL, int pi,
            int nRates, num_t mu, num_t var, num_t dT, num_t beta, num_t alpha, num_t eps,
            num_t minSumJump, int nUnits)
 :
+    d_id(s_id),
     d_nUnits(nUnits),
     d_ccm(ccm),
     d_cpm(cpm),
@@ -29,9 +32,12 @@ Init::Init(num_t ccm, num_t cpm, num_t C, int nStates, num_t dL, int pi,
         throw("Number of units should be either 2 or 3");
     
     initialise();
+    ++s_id;
 }
 
-Init::Init(string const &line) 
+Init::Init(string const &line)
+:
+    d_id(s_id)
 {
     // Split line into words seperated by spaces and put it in vector
     vector<string> args;
@@ -69,4 +75,6 @@ Init::Init(string const &line)
 
     // Initialise the rest
     initialise();
+
+    ++s_id;
 }

### 2018-06-22 src/init/init.h

@@ -13,6 +13,10 @@ class Init
 {
     friend class MDP;
 
+    // ID 
+    static size_t   s_id;
+    size_t const    d_id;
+
     // Number of units
     int d_nUnits;
 

### 2018-06-22 src/mdp/direct_cost.cpp

@@ -3,11 +3,10 @@
 num_t MDP::direct_cost(int state1, int state2, int mainAction1, int mainAction2) const
 {
     // First check if there is any maintenance action
-    int sumAction = mainAction1 + mainAction2;
-    if (sumAction == 0)
+    if (mainAction1 + mainAction2 == 0)
         return 0;
 
-    return d_C + direct_costs(state1, mainAction1) + direct_costs(state2, mainAction2);
+    return (d_C + direct_costs(state1, mainAction1) + direct_costs(state2, mainAction2));
 }
 
 num_t MDP::direct_cost(int state1, int state2, int state3, 
@@ -18,5 +17,5 @@ num_t MDP::direct_cost(int state1, int state2, int state3,
     if (sumAction == 0)
         return 0;
 
-    return d_C + direct_costs(state1, mainAction1) + direct_costs(state2, mainAction2) + direct_costs(state3, mainAction3);
+    return (d_C + direct_costs(state1, mainAction1) + direct_costs(state2, mainAction2) + direct_costs(state3, mainAction3));
 }

### 2018-06-22 src/mdp/mdp.cpp

@@ -1,10 +1,8 @@
 #include "mdp.ih"
 
-size_t MDP::s_id = 0;
-
 MDP::MDP(Init const &init, bool jumpHeuristic)
 :
-    d_id(s_id),
+    d_id(init.d_id),
     d_valueA(vector<num_t>(init.d_nStates * init.d_nStates, 0)),
     d_valueB(vector<num_t>(init.d_nStates * init.d_nStates, 0)),
     d_expectedCost(vector<num_t>(init.d_nStates * init.d_nStates, 0)),
@@ -32,8 +30,6 @@ MDP::MDP(Init const &init, bool jumpHeuristic)
 {
     if (d_pi > 2 * (d_nRates-1))
         throw string("Error: pi should be feasible for units");
-
-    ++s_id;
 }
 
 

### 2018-06-22 src/mdp/mdp.h

@@ -10,7 +10,6 @@
 
 class MDP
 {
-    static size_t   s_id;
     size_t const    d_id;
     
     // Values

### 2018-06-22 src/mdp/value_costs.cpp

@@ -22,7 +22,7 @@ num_t MDP::value_costs(int state1, int state2, int idxState)
             int s2 = (adx2 == 0) ? state2 : 0;
 
             num_t const tmpVal = direct_cost(state1, state2, adx1, adx2) + d_expectedCost[hash(s1, s2)];
-            if (tmpVal < bestVal)
+            if (tmpVal < bestVal - gg_eps)
             {
                 bestVal = tmpVal;
                 bestMain1 = adx1;

### 2018-06-22 src/mdp/value_costs_heuristic.cpp

@@ -33,7 +33,7 @@ num_t MDP::value_costs_heuristic2(int state1, int state2, int idxState)
             int s2 = (adx2 == 0) ? state2 : 0;
 
             num_t const tmpVal = direct_cost(state1, state2, adx1, adx2) + d_expectedCostHeuristic[hash(s1, s2)];
-            if (tmpVal < bestVal)
+            if (tmpVal < bestVal - gg_eps)
             {
                 bestVal = tmpVal;
                 bestMain1 = adx1;

### 2018-06-22 src/run.cpp

@@ -14,12 +14,12 @@ try
 {
     if (argc == 2)
     {
-        solve_from_file();
+        solve_from_file(true);
     } else
     {
-        Init init{1,        // ccm
-                  0.2,      // cpm
-                  0.5,      // C
+        Init init{11.0,        // ccm
+                  5.0,      // cpm
+                  4.0,      // C
                   21,       // nStates
                   1,        // dL
                   20,       // pi
@@ -29,7 +29,7 @@ try
                   1,        // dT
                   0.1,      // beta
                   1.5,      // alpha
-                  0.000005, // eps
+                  0.0000005, // eps
                   0.99,     // minProb
                   2         // nUnits
                  };

### 2018-06-22 src/solve_experiments.h

@@ -4,21 +4,20 @@
 #include <iostream>
 #include <fstream>
 
+#include <thread>
+
 #include "init/init.h"
 #include "mdp/mdp.h"
 
-void solve_experiment(Init const &init)
+void solve_experiment(Init const init)
 {
     // Initialise MDP
-    init.printAllInfo();
     MDP mdp{init, false};
     mdp.solve();
-    mdp.print_all_info();
     mdp.write_solution();
-    std::cout << "Solved experiment " << mdp.get_id() << std::endl;
 }
 
-void solve_from_file()
+void solve_from_file(bool const multi_thread)
 {
     // Read experiment file
     std::string line;
@@ -40,7 +39,20 @@ void solve_from_file()
     }
 
     // Solve all experiments in different threads
-    for (Init const &init: init_vec)
-        solve_experiment(init);
+    if (multi_thread)
+    {
+        std::vector<std::thread> thread_vec;
+        for (Init const &init: init_vec)
+            thread_vec.push_back(std::thread(solve_experiment, init));
+        
+        // Wait for all threads to join
+        for (auto &thread: thread_vec)
+            thread.join();
+
+    } else 
+    {
+        for (Init const &init: init_vec)
+            solve_experiment(init);
+    }
 }
 #endif
